---
format:
    html
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7.5,
  fig.height = 4
)

# update 'asis' chunk to allow inline code
knitr::knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knitr::knit_child(text = options$code)
})
```

```{r}
#| label: setup-demos-unstratified
#| message: false
# library(fxtas)
devtools::load_all()
library(pander)
library(table1)
library(dplyr)
library(tidyr)
library(magrittr)
library(arsenal)
```

```{r, include=FALSE}
# scid vars
scid_vars = c(
    "Bipolar I Disorder (MD01), Lifetime",                            
    "Bipolar I Disorder (MD01), Current",
    "Bipolar II Disorder (MD02), Lifetime",         
    "Bipolar II Disorder (MD02), Current",         
    "Other Bipolar Disorder (MD03), Current",         
    "Other Bipolar Disorder (MD03), Lifetime",         
    "Major Depressive Disorder (MD04), Current",         
    "Major Depressive Disorder (MD04), Lifetime",         
    "Mood Disorder Due to GMC (MD07), Current",         
    "Mood Disorder Due to GMC (MD07), Lifetime",         
    "Substance-Induced Mood Dis. (MD08), Current",         
    "Substance-Induced Mood Dis. (MD08), Lifetime",         
    "Primary Psychotic Symptoms (PS01), Current",         
    "Primary Psychotic Symptoms (PS01), Lifetime",
    "Dysthymic Disorder (MD05)",         
    "Depressive Disorder NOS (MD06), Lifetime",         
    "Depressive Disorder NOS (MD06) Current",         
    "Alcohol (SUD17), Lifetime",         
    "Alcohol (SUD17) Current",         
    "Sedative-Hypnotic-Anxiolytic (SUD18), Lifetime",         
    "Sedative-Hypnotic-Anxiolytic (SUD18), Current",         
    "Cannabis (SUD19), Lifetime",         
    "Cannabis (SUD19),Current",         
    "Stimulants (SUD20), Lifetime",         
    "Stimulants (SUD20) Current",         
    "Opiod (SUD21), Lifetime",         
    "Cocaine (SUD22), Lifetime",         
    "Opiod (SUD21), Current",         
    "Hallucinogenics/ PCP (SUD23), Lifetime",         
    "Cocaine (SUD22) Current",         
    "Hallucinogenics/ PCP (SUD23), Current",         
    "Poly Drug (SUD24), Lifetime",         
    "Poly Drug (SUD24), Current",         
    "Other (SUD25), Lifetime",         
    "Other (SUD25), Current",         
    "Panic Disorder (ANX26), Lifetime",         
    "Panic Disorder (ANX26), Current",         
    "Agoraphobia without Panic (ANX27), Lifetime",         
    "Agoraphobia without Panic (ANX27), Current",         
    "Social Phobia (ANX28), Lifetime",         
    "Social Phobia (ANX28), Current",         
    "Specific Phobia (ANX29), Lifetime",         
    "Specific Phobia (ANX29), Current",         
    "Obsessive Compulsive (ANX30), Lifetime",         
    "Obsessive Compulsive (ANX30), Current",         
    "Posttraumatic Stress (ANX31), Lifetime",         
    "Posttraumatic Stress (ANX31), Current",         
    "Generalized Anxiety (ANX32), Current Only",         
    "Anxiety Due to GMC (ANX33), Lifetime",         
    "Anxiety Due to GMC (ANX33), Current",         
    "Substance-Induced Anxiety (ANX34), Lifetime",         
    "Substance-Induced Anxiety (ANX34), Current",         
    "Anxiety Disorder NOS (ANX35), Lifetime",         
    "Anxiety Disorder NOS (ANX35), Current",         
    "Somatization Disorder (SOM36)",         
    "Pain Disorder (SOM37)",         
    "Undifferentiated Somatoform (SOM38)",         
    "Body Dysmorphic (SOM40)",         
    "Hypochondriasis (SOM39)"
  )

# scid vars by domain
scid_md_vars <- scid_vars[grepl("\\(MD", scid_vars)]
scid_sud_vars <- scid_vars[grepl("\\(SUD", scid_vars)]
scid_anx_vars <- scid_vars[grepl("\\(ANX", scid_vars)]
scid_somatic_vars <- scid_vars[grepl("\\(SOM", scid_vars)]
scid_psych_vars <- scid_vars[grepl("\\(PS", scid_vars)]

# scid vars by time (lifetime vs current)
lifetime_vars <- scid_vars[grepl("Lifetime", scid_vars)]
current_vars <- scid_vars[grepl("Current", scid_vars)]
# remove Current Only
current_vars <- current_vars[!grepl("Current Only", current_vars)]

# subset gp34 to scid vars
data("gp34")
scid_dat <- gp34 |>
  dplyr::select(`FXS ID`, `Event Name`, all_of(scid_vars))

# convert to long format
scid_lifetime_long <- scid_dat |>
  dplyr::select(`FXS ID`, `Event Name`, all_of(lifetime_vars)) |>
  pivot_longer(
    cols = all_of(lifetime_vars),
    names_to = "Variable",
    values_to = "Lifetime"
  ) |>
  mutate(
    Variable = gsub("Lifetime", "", Variable)
  ) |>
  mutate(
    Variable = gsub("\\,", "", Variable)
  ) |>
  mutate(
    Variable = trimws(Variable)
  ) |>
  mutate(
    Lifetime = factor(Lifetime, levels = c("Absent", "Sub-Threshold", "Threshold"))
  )
  

scid_current_long <- scid_dat |>
  dplyr::select(`FXS ID`, `Event Name`, all_of(current_vars)) |>
  pivot_longer(
    cols = all_of(current_vars),
    names_to = "Variable",
    values_to = "Current"
  ) |>
  mutate(
    Variable = gsub("Current", "", Variable)
  ) |>
  mutate(
    Variable = gsub("\\,", "", Variable)
  ) |>
  mutate(
    Variable = trimws(Variable)
  ) |>
  mutate(
    Current = factor(Current, levels = c("Absent", "Present"))
  )

lifetime_vars_clean <- unique(scid_lifetime_long$Variable)
current_vars_clean <- unique(scid_current_long$Variable)

lifetime_vars_clean[!(lifetime_vars_clean %in% current_vars_clean)]
current_vars_clean[!(current_vars_clean %in% lifetime_vars_clean)]

scid_long_vars <- c(
  "Bipolar I Disorder (MD01)", "Bipolar II Disorder (MD02)", 
  "Other Bipolar Disorder (MD03)", "Major Depressive Disorder (MD04)", 
  "Depressive Disorder NOS (MD06)", "Mood Disorder Due to GMC (MD07)", 
  "Substance-Induced Mood Dis. (MD08)", "Primary Psychotic Symptoms (PS01)",  
  "Alcohol (SUD17)", "Sedative-Hypnotic-Anxiolytic (SUD18)", "Cannabis (SUD19)", 
  "Stimulants (SUD20)", "Opiod (SUD21)", "Cocaine (SUD22)", 
  "Hallucinogenics/ PCP (SUD23)", "Poly Drug (SUD24)", "Other (SUD25)", 
  "Panic Disorder (ANX26)", "Agoraphobia without Panic (ANX27)", 
  "Social Phobia (ANX28)", "Specific Phobia (ANX29)", "Obsessive Compulsive (ANX30)", 
  "Posttraumatic Stress (ANX31)", "Anxiety Due to GMC (ANX33)", 
  "Substance-Induced Anxiety (ANX34)", "Anxiety Disorder NOS (ANX35)"
)

scid_long <- merge(scid_current_long, scid_lifetime_long, 
                   by = c("FXS ID", "Event Name", "Variable"), sort = FALSE) |>
  mutate(
    Variable = factor(
      Variable, levels = scid_long_vars
    )
  ) |> 
  as_tibble()


# test scid domains
scid_domains <- create_scid_domains(gp34 |> get_visit1())
```

# SCID: Current vs Lifetime

For the disorders with current vs lifetime variables, examine the relationship between them. If `current` is present, is `lifetime` always sub-threshold or threshold?  

```{r, results='asis'}
summary(
  arsenal::tableby(
    na.action = na.tableby(lhs = TRUE),
    includeNA(Lifetime) ~ includeNA(Current),
    strata = Variable,
    data = scid_long,
    test = FALSE,
    total = FALSE
  )
)
```

Occasionally, current = "Present" and lifetime = "Absent":

```{r}
scid_long |> filter(Current == "Present", Lifetime %in% c("Absent", NA))

```

double-checking the first case:

```{R}
gp34 |> 
  filter(`FXS ID` == "104279-100") |> 
  select(`Event Name`, 
         `Anxiety Disorder NOS (ANX35), Lifetime`, 
         `Anxiety Disorder NOS (ANX35), Current`)
```

Missingness issues:

```{r}
scid_long |> filter(!is.na(Current), is.na(Lifetime))
```


# SCID: Domains

Using only `lifetime` variables or variables without `current`/`lifetime` distinction, collapse the SCID variables into 5 domains: Mood Disorders, Anxiety, Substance Use Disorder, Somatization, and Psychotic.

```{r, results='asis'}
summary(
  arsenal::tableby(
    `FX**` ~ `SCID: Mood Disorders` + `SCID: Substance Use` + 
      `SCID: Anxiety` + `SCID: Somatization` + `SCID: Psychotic`,
    data = scid_domains,
    test = FALSE
  )
)
```
