---
format:
    html
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7.5,
  fig.height = 4
)

# update 'asis' chunk to allow inline code
knitr::knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knitr::knit_child(text = options$code)
})
```

```{r}
#| label: setup-demos-unstratified
#| message: false
library(fxtas)
# devtools::load_all()
library(pander)
library(table1)
library(dplyr)
library(tidyr)
library(magrittr)
library(arsenal)

NA.label1 = "Missing"
column_var = "FX3*"

data("visit1")

visit1_cgg_complete = 
  visit1 |> 
  filter(
    CGG < 200,
    !is.na(`FX*`)
    )
```

# MRI Variables

Create MRI cerebral and cerebellum domains.

<ul>
  <li>Cerebral variables:
    <ul>
      <li>Cerebral Atrophy</li> 
      <li>Cerebral WM Hyperintensity</li> 
      <li>Pons-WM Hyperintensity</li>
      <li>Sub-Insular WM Hyperintensity</li> 
      <li>Periventricular WM Hyperintensity</li>
      <li>Splenium (CC)-WM Hyperintensity</li>
    </ul>
  </li>
  <li>Cerebellar variables:
    <ul>
      <li>Cerebellar Atrophy</li>
      <li>Cerebellar WM Hyperintensity</li>
      <li>MCP-WM Hyperintensity</li>
    </ul>
  </li>
</ul>


## Cerebral Variables 

```{r}
vars <- c(
  "Cerebral Atrophy", "Cerebral WM Hyperintensity", "Pons-WM Hyperintensity",
  "Sub-Insular WM Hyperintensity", "Periventricular WM Hyperintensity",
  "Splenium (CC)-WM Hyperintensity", "MRI: Cerebral"
)

table1(na.is.category = FALSE, 
  NA.label = NA.label1,
  
  paste(
    "~", 
    paste(
      formulaic::add.backtick(vars), 
      collapse = " + "), 
    "|",
    column_var |> formulaic::add.backtick()) |> 
    formula(),
  # render.missing=c("Empty in RedCap"="FREQ (PCT%)"),
  data = visit1_cgg_complete)
```

## Cerebellar Variables

```{r}
vars <- c(
  "Cerebellar Atrophy", "Cerebellar WM Hyperintensity",
  "MCP-WM Hyperintensity", "MRI: Cerebellum"
)

table1(na.is.category = FALSE, 
  NA.label = NA.label1,
  
  paste(
    "~", 
    paste(
      formulaic::add.backtick(vars), 
      collapse = " + "), 
    "|",
    column_var |> formulaic::add.backtick()) |> 
    formula(),
  # render.missing=c("Empty in RedCap"="FREQ (PCT%)"),
  data = visit1_cgg_complete)
```
