```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7.5,
  fig.height = 4
)
```

```{r}
#| label: setup-demos-unstratified
#| message: false
library(fxtas)
# devtools::load_all()
library(pander)
library(table1)
library(dplyr)
library(magrittr)

# dealing with misssing values:
NA.label1 = "Missing"
show_missingness_reasons = FALSE
footnote1 = 'Codes `777`, `888`, and `999` are counted as "Missing" (NA).'

column_var = "Recruited in study phase"
column_var = "FX3*"
column_var = c("Gender", "FX3*")

visit1_cgg_complete = 
  trax_gp34_v1 |> 
  filter(
    CGG < 200,
    !is.na(`FX*`)
  )

phase <- ifelse(column_var == "Recruited in study phase", TRUE, FALSE)
cgg <- ifelse(column_var == "FX**", TRUE, FALSE)
```

## First-visit characteristics by Sex and CGG repeats level

Here, we review the data from the first visits by each participant, stratified by CGG (`r unique(visit1_cgg_complete$"FX3*")`) 
and Sex (`r unique(visit1_cgg_complete$Gender)`)

### Demographics

```{r}
#| label: "tbl-demographics"
#| tbl-cap: "Demographics by CGG repeats level"
vars = c(
  "Study",
  "Age at first visit",
  "# visits",
  # column_var,
  "Gender",
  "Primary Ethnicity",
  "Primary Race"
  
)

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = visit1_cgg_complete)
```

### APOE

```{r}
#| tbl-cap: "APOE by CGG repeats level"
#| label: "tbl-apoe"
# vars = grep("^ApoE", names(trax_gp34_all), value = TRUE)
vars = "ApoE"
table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  x = stratified_formula(vars, column_var),
  # render.missing=c("Empty in RedCap"="FREQ (PCT%)"),
  data = visit1_cgg_complete)
```

### Tremors

```{r}
vars = c(
  
  "Intention tremor",
  "Resting tremor",
  "Postural tremor",
  "Intermittent tremor",
  "Any tremor (excluding Head Tremor)"
  # "Tremor: Age of onset"
  # if(show_missingness_reasons) "Tremor: Age of onset: missingness reasons"
)
table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       render.continuous=c(
         .="N", 
         .="Mean (SD)", 
         .="Median [Min, Max]"),
       data = visit1_cgg_complete)

```

<br>

```{r}
#| tbl-cap: "Age of onset for those with 'Some tremors recorded'."
#| label: "tbl-tremor-ao"
#| 
vars = c(
         "Tremor: Age of onset",
         # "Tremor: Age of onset: missingness reasons",
         NULL
         )


table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = visit1_cgg_complete |> 
         filter(`Any tremor (excluding Head Tremor)` == "Some tremors recorded"))
```

<br>

### Head tremors

```{r}
#| tbl-cap: Head tremors
#| label: tbl-head-tremors
vars = c(
  "Head tremor"
  # "Head Tremor: Age of onset"
  
)

footnote_HT <- paste(
  "Head tremor was not collected in GP3 or Trax Phase 1.",
  "Summaries provided are for GP4 + Trax Phase 2 only."
)

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = visit1_cgg_complete |> filter(!(Study %in% c("Trax Phase 1", "GP3"))),
  footnote = footnote_HT)

```

<br>

```{r}
#| tbl-cap: "Age of onset for those with a head tremor recorded."
#| label: "tbl-head-tremor-AO"
#| eval: false
vars = c(
  "Head Tremor: Age of onset",
  if(show_missingness_reasons) "Head Tremor: Age of onset: missingness reasons"
  
)

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = visit1_cgg_complete |> filter(`Head tremor` == "Yes"))

```

### Ataxia

```{r}
#| label: tbl-ataxia-by-cgg
#| tbl-cap: "Ataxia by CGG status"
vars = c(
  "Ataxia",
  "Ataxia: severity",
  "Ataxia: severity*"
  # "Ataxia: Age of onset",
  # "Ataxia: Age of onset: missingness reasons"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = visit1_cgg_complete)

```

<br>

```{r}
#| label: tbl-ataxia-AO-by-cgg
#| tbl-cap: "Ataxia age of onset by CGG level"
vars = c(
  # "Ataxia",
  # "Ataxia: severity",
  # "Ataxia: severity*"
  "Ataxia: Age of onset"
  # "Ataxia: Age of onset: missingness reasons"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = visit1_cgg_complete |> filter(Ataxia == "Yes"))
```

<br>

### Parkinsons

```{r}
#| label: tbl-parkinsons-by-cgg
#| tbl-cap: "Parkinsons by CGG status"
vars = grep("Parkinson", value = TRUE, names(trax_gp34_all))

if(!show_missingness_reasons) 
{
  vars = vars |> remove_missingness_reason_vars()
}

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       data = visit1_cgg_complete)

```

### FXTAS Stage

```{r}
#| label: tbl-fxtas-stage-by-cgg
#| tbl-cap: "FXTAS Stage by CGG level"
vars = c(
  "FXTAS Stage (0-5)*",
  if(show_missingness_reasons) "FXTAS Stage (0-5)*: missingness reasons"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       footnote = "For categorical analysis, we merged 2.5 into 2, 3.5 into 4, and 4.5 into 5.",
       data = visit1_cgg_complete)

# trax_gp34_all |> 
#   filter(`FXTAS Stage (0-5)` %in% c(NA, "No Response (999)", "Missing (empty in RedCap)")) |> 
#   select(`FXS ID`, `Event Name`, `Visit Date`) |> 
#   arrange() |> 
#   readr::write_csv("inst/extdata/missing_FXTAS_stage.csv")

```

### BDS2

```{r}

vars = c("BDS-2 Total Score", "BDS-2 Total Score*")


table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       footnote = footnote1,
       data = visit1_cgg_complete)

```

### MMSE

```{r}

# vars = grep(
#   "MMSE Total Score", 
#   value = TRUE, 
#   names(trax_gp34_all), 
#   fixed = TRUE)

vars = c(
  "MMSE Total Score", 
  # "MMSE Total Score: missingness reasons", 
  "MMSE Total Score*")

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       footnote = footnote1,
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = visit1_cgg_complete)

```

### SCID

```{r}

vars = scid_vars_for_table1()

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = visit1_cgg_complete)

```



### MRI

```{r}
#| label: tab-mri
#| tbl-cap: "MRI variables by Sex and CGG level"
vars = c(
  "Cerebral Atrophy",
  "Cerebellar Atrophy",
  "Cerebral WM Hyperintensity",
  "Cerebellar WM Hyperintensity",
  "MCP-WM Hyperintensity",
  "Pons-WM Hyperintensity",
  "Sub-Insular WM Hyperintensity",
  "Periventricular WM Hyperintensity",
  "Splenium (CC)-WM Hyperintensity",
  "Genu (CC)-WM Hyperintensity",
  "Corpus Callosum-Thickness"
)

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  data = visit1_cgg_complete)

```

### WAIS

```{r}

vars = c(
  # "Verbal: IQ Score",
  "Verbal Comprehension: Composite Score (VCI)",
  
  "Perceptual Reasoning: Composite Score (PRI)",
  "Working Memory: Composite Score (WMI)", 
  "Processing Speed: Composite Score (PSI)",
  
  # "Full Scale: IQ Score",
  "Full Scale: Composite Score (FSIQ)"
)

if(show_missingness_reasons)
{
  vars = vars |> 
    sapply(
      F = function(x) 
        grep(
          value =  TRUE, 
          fixed = TRUE, 
          pattern = x, 
          x = names(visit1_cgg_complete))) |> as.vector()
}

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min, Max]"),
  data = visit1_cgg_complete)

```

### CANTAB

For now, we used cutoffs taken from [@talebi2020cambridge].

```{r}

# table1(na.is.category = FALSE, overall = FALSE,
#   NA.label = NA.label1,
#   stratified_formula("Drugs used", column_var),
#   data = visit1_cgg_complete)


cantab_vars = c(
  "SWM Between errors",
  "SWM Between errors*",
  "SST Median correct RT on GO trials",
  "RVP A signal detection",
  "RVP A signal detection*",
  "OTS Problems solved on first choice",
  "PAL Total errors (adjusted)",
  "PAL Total errors (adjusted)*",
  "RTI Five-choice movement time",
  "RTI Five-choice movement time*") 
# |> 
#   sapply(F = function(x) grep(value =  TRUE, fixed = TRUE, pattern = x, x = names(visit1_cgg_complete))) |> 
# as.vector()

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = "Missing (empty in RedCap)",
       stratified_formula(cantab_vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = visit1_cgg_complete)
```

### Thyroid and other autoimmune diseases

```{r}

vars = c(
  "Hypothyroid",
  "Hyperthyroid",
  "Thyroid problems",
  "Lupus",
  "Rheumatoid arthritis",
  "Multiple Sclerosis: Workup",
  "ANA positive",
  "Sjogrens Syndrome",
  "Raynauds Syndrome",
  "Pulmonary Fibrosis",
  "Any Autoimmune"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       data = visit1_cgg_complete)
```
