Here we compared the longitudinal models to the cross-sectional models (estimated using each participant's earliest visit):

- Males only (@fig-males):
  * first visits only (@fig-males-1)
  * all visits (@fig-males-2)

- Females only (@fig-females):
  * first visits only (@fig-females-1)
  * all visits (@fig-females-2)

We also rearranged these results by subset type:

- first visits only (@fig-first-only)
- all visits (@fig-all-visits)


::: landscape

```{r}
fig_females_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "females",
  output_folder = "output/gp34_longitudinal")

fig_females_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "females",
  output_folder = "output/gp34_multivisit_only")


fig_males_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "males",
  output_folder = "output/gp34_longitudinal")

fig_males_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "males",
  output_folder = "output/gp34_multivisit_only")


fig_both_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "gp34_longitudinal",
  output_folder = "output/gp34_longitudinal")

fig_both_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "gp34_multivisit_only",
  output_folder = "output/gp34_multivisit_only")

```

### Males {.appendix}

@fig-males shows estimated event sequences for males, depending on which subset of the data we use to fit the model.

```{r}
#| label: "fig-males"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences for males, by data subset"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "First visits only"
#|   - "All visits" 

fig_males_first |> print()

fig_males_all |> print()

# fig_males_mv_only |> print()

```

### Females {.appendix}

```{r}
#| label: "fig-females"
#| fig-cap: "Estimated event sequences for females, by data subset"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "First visits only"
#|   - "All visits" 

fig_females_first |> print()

fig_females_all |> print()

# fig_females_mv_only |> print()

```

### All patients, all visits  {.appendix}

@fig-all-visits shows the event sequences that we estimated from each data subset. Note that clinical events which never occur in a dataset cannot be placed into the event sequence, so the set of modeled events varies by data subset.

We note that Parkinsonian features mostly appear after FXTAS stage 4 for females, whereas they occur before stage 4 for males.

Similarly, clinically-elevated SCL90 values appear before FXTAS stage 1 for females, whereas they occur after stage 3 for males.

{{< pagebreak >}}

```{r}
#| label: "fig-all-visits"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences - all visits"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "Males only" 
#|   - "Females only"

# fig_both_all |> print()
fig_males_all |> print()
fig_females_all |> print()

```

{{< pagebreak >}}

```{r}
#| label: fig-stage-by-age-longitudinal
#| fig-cap: "ML stage by age, longitudinal data"
#| include: true

patient_data1 = 
  trax_gp34_all |>
  filter(
    CGG < 200, 
    CGG >= 55)

results_both_all = extract_results_from_pickle(
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "gp34_longitudinal",
  output_folder = "output/gp34_longitudinal")

stages = 
  results_both_all$subtype_and_stage_table |> 
  mutate(
    age = patient_data1$`Age at visit`,
    id = patient_data1$`FXS ID`
  ) |> 
  arrange(id, age)

library(ggplot2)
stages |> graph_stage_by_age()


```

:::
