---
title: "Modeling longitudinal data"
knitr:
  opts_chunk: 
    collapse: true
    comment: "#>" 
    R.options:  
      dev: svg

project:
  execute-dir: project
---

```{=html}
<style>
.quarto-figure-center > figure {
text-align: center;
}
</style>
```

# Methods

Next, we fit the Ordinal SuStaIn model [@young2021ordinal] on nine subsets of the longitudinal GP+GP4 data:

- Males only (@fig-males):
  * first visits only (@fig-males-1)
  * all visits (@fig-males-2)
  * multivisit patients only (@fig-males-3)

- Females only (@fig-females):
  * first visits only (@fig-females-1)
  * all visits (@fig-females-2)
  * multivisit patients only (@fig-females-3)
  
- Unstratified (@fig-both-genders):
  * first visits only (@fig-both-genders-1)
  * all visits (@fig-both-genders-2)
  * multivisit patients only (@fig-both-genders-3)
  
We also rearranged these results by subset type:

- first visits only (@fig-first-only)
- all visits (@fig-all-visits)
- multivisit patients only (@fig-mv-only)



```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # fig.width = 7, 
  # fig.height = 10,
  include = TRUE
)

devtools::load_all()
# library(fxtas)
library(tidyverse)
library(pander)
```

```{r}
#| label: extract-figs


fig_females_first = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "females",
  output_folder = "output/output.fixed_CV")

fig_females_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "females",
  output_folder = "output/gp34_longitudinal")

fig_females_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "females",
  output_folder = "output/gp34_multivisit_only")


fig_males_first = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "males",
  output_folder = "output/output.fixed_CV")

fig_males_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "males",
  output_folder = "output/gp34_longitudinal")

fig_males_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "males",
  output_folder = "output/gp34_multivisit_only")

fig_both_first = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "sample_data",
  output_folder = "output/output.fixed_CV")

fig_both_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "gp34_longitudinal",
  output_folder = "output/gp34_longitudinal")

fig_both_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "gp34_multivisit_only",
  output_folder = "output/gp34_multivisit_only")

```


# Results

## Males

@fig-males shows estimated event sequences for males, depending on which subset of the data we use to fit the model.

```{r}
#| label: "fig-males"
#| column: page
#| fig-height: !expr pvd_height
#| fig-align: center
#| fig-cap: "Estimated event sequences for males, by data subset"
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "First visits only"
#|   - "All visits" 
#|   - "Multivisit patients only"

fig_males_first |> print()

fig_males_all |> print()

fig_males_mv_only |> print()

```

## Females

```{r}
#| label: "fig-females"
#| fig-cap: "Estimated event sequences for females, by data subset"
#| column: page
#| fig-height: !expr pvd_height
#| fig-align: center
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "First visits only"
#|   - "All visits" 
#|   - "Multivisit patients only"

fig_females_first |> print()

fig_females_all |> print()

fig_females_mv_only |> print()

```


## Unstratified, by data subset

```{r}
#| label: "fig-both-genders"
#| column: page
#| fig-height: !expr pvd_height
#| fig-align: center
#| fig-cap: "Estimated event sequences, unstratified, by data subset"
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "First visits only"
#|   - "All visits" 
#|   - "Multivisit patients only"

fig_both_first |> print()
fig_both_all |> print()
fig_both_mv_only |> print()

```

## All patients, first visits only

```{r}
#| label: "fig-first-only"
#| column: page
#| fig-height: !expr pvd_height
#| fig-align: center
#| fig-cap: "Estimated event sequences - first visits only"
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "Unstratified"
#|   - "Males only" 
#|   - "Females only"

fig_both_first |> print()
fig_males_first |> print()
fig_females_first |> print()


```

## All patients, all visits

@fig-all-visits shows the event sequences that we estimated from each data subset. Note that clinical events which never occur in a dataset cannot be placed into the event sequence, so the set of modeled events varies by data subset.

```{r}
#| label: "fig-all-visits"
#| column: page
#| fig-height: !expr pvd_height
#| fig-align: center
#| fig-cap: "Estimated event sequences - all visits"
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "Unstratified"
#|   - "Males only" 
#|   - "Females only"

fig_both_all |> print()
fig_males_all |> print()
fig_females_all |> print()

```

## Multivisit patients only, all visits

@fig-mv-only shows the event sequences that we estimated using only the data from patients with more than one visit.

```{r}
#| label: "fig-mv-only"
#| column: page
#| fig-height: !expr pvd_height
#| fig-align: center
#| fig-cap: "Estimated event sequences - multivisit patients only"
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "Unstratified"
#|   - "Males only" 
#|   - "Females only"

fig_both_mv_only |> print()
fig_males_mv_only |> print()
fig_females_mv_only |> print()

```

# References {.unnumbered}

::: {#refs}
:::
