---
title: "Ordinal SuStaIn"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{SuStaIn-ModelFitting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 7,
  include = TRUE
)
```

```{r setup}
library(fxtas)
library(tidyverse)
library(reticulate)
```


# Import data

```{r}

tremors = c(
  "Intention tremor" = 1,
  "Resting tremor" = 1,
  "Postural tremor" = 1,
  "Intermittent tremor" = 1
  # "Any tremor" = 1
)

mri_vars = c(
  "Cerebral Atrophy",
  "Cerebellar Atrophy",
  "Cerebral WM Hyperintensity",
  "Cerebellar WM Hyperintensity",
  "MCP-WM Hyperintensity",
  "Pons-WM Hyperintensity",
  "Sub-Insular WM Hyperintensity",
  "Periventricular WM Hyperintensity",
  "Splenium (CC)-WM Hyperintensity",
  "Genu (CC)-WM Hyperintensity",
  "Corpus Callosum-Thickness"
)

SuStaInLabels = 
  biomarkers =
  # mri_vars
  names(tremors)

df = 
  visit1 |> 
  select(
    `FXS ID`,
    `FX*`,
    all_of(biomarkers)) |> 
  filter(
    !is.na(`FX*`)) |> 
  mutate(
    across(all_of(biomarkers), ~ as.numeric(.x) - 1)
  ) |> 
  na.omit()

```

# Set up data for new version of SuStaIn
The objective here is two create to variables:

* `p_NL` -- a subject x region matrix indicating the p that a region is normal for a subject.
* `p_score` -- a subject x (region * score) matrix indicating p that a region is a given score.

Ps should be non-zero. 

```{r}

ModelScores = DataScores = 
  df[,biomarkers] |> compute_score_levels()

control_data = 
  df |> 
  filter(`FX*` == "CGG < 55") |> 
  select(all_of(biomarkers))

patient_data = 
  df |> 
  filter(`FX*` == "CGG >= 55") |> 
  select(all_of(biomarkers))

prob_correct = 
  control_data |> 
  compute_prob_correct(
    biomarkers = biomarkers,
    DataScores = DataScores)

# prob_correct[] = .95

prob_dist = compute_prob_dist(
  ModelScores,
  DataScores,
  biomarkers,
  prob_correct
)

prob_nl_dist = prob_dist[1, , ]
prob_score_dist = prob_dist[-1, , ]

```

```{r}

prob_score0 = compute_prob_scores(
  df,
  biomarkers,
  ModelScores,
  DataScores,
  prob_dist
)

prob_nl = prob_score0[,,1]
prob_score = prob_score0[,,-1, drop = FALSE]

```

```{r "score_vals"}

score_vals = matrix(
  ModelScores[-1] |> as.numeric(),
  byrow = TRUE,
  nrow = length(biomarkers),
  ncol = length(ModelScores) - 1)

```

# Run SuStaIn model

```{r "set run parameters"}

N_startpoints = 10L
N_S_max = 3L
N_iterations_MCMC = 1e5L
dataset_name = 'sample_data'
output_folder = "output.FXTAS2" |> fs::dir_create()
print(output_folder)
```

```{python "initialize-model"}

import pandas
import numpy as np
import numpy.matlib
import matplotlib.pyplot as plt
import seaborn as sns
import sys
import scipy
from glob import glob
import pySuStaIn
import os

os.getcwd()

sustain_input = pySuStaIn.OrdinalSustain(r.prob_nl,
r.prob_score,
r.score_vals,
r.SuStaInLabels,
r.N_startpoints,
r.N_S_max, 
r.N_iterations_MCMC, 
r.output_folder, 
r.dataset_name, 
False,
1,
True)

```

# fitting model


```{python "fit-model", fig.width = 8, fig.height = 9}

[samples_sequence, samples_f, 
ml_subtype, prob_ml_subtype, ml_stage, 
prob_ml_stage, prob_subtype_stage, 
samples_likelihood] = sustain_input.run_sustain_algorithm(plot=True)

plt.show()
```

# calling plot function explicitly

```{python,  fig.width = 8, fig.height = 9}
sustain_input._plot_sustain_model(samples_sequence = samples_sequence,
samples_f = samples_f,
n_samples = r.df.shape[0],
biomarker_labels = r.SuStaInLabels)
_ = plt.suptitle('SuStaIn output')

```

# graph likelihoods in R

```{r, fig.width = 8, fig.height = 9}
py$samples_likelihood |>
  graph_likelihoods_v2(alpha = 0.5)

```

### trying `plt.show()`

```{python, fig.width = 8, fig.height = 9}
plt.show()
```

