---
title: "Ordinal SuStaIn"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{SuStaIn-ModelFitting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  fig.width = 4, 
  fig.height = 7,
  include = FALSE
)
```

```{r setup}
library(fxtas)
library(tidyverse)
library(reticulate)
```


# Import data

```{r}

tremors = c(
  "Intention tremor" = 1,
  "Resting tremor" = 1,
  "Postural tremor" = 1,
  "Intermittent tremor" = 1,
  "Any tremor" = 1
)

mri_vars = c(
  "Cerebral Atrophy",
  "Cerebellar Atrophy",
  "Cerebral WM Hyperintensity",
  "Cerebellar WM Hyperintensity",
  "MCP-WM Hyperintensity",
  "Pons-WM Hyperintensity",
  "Sub-Insular WM Hyperintensity",
  "Periventricular WM Hyperintensity",
  "Splenium (CC)-WM Hyperintensity",
  "Genu (CC)-WM Hyperintensity",
  "Corpus Callosum-Thickness"
)

SuStaInLabels = 
  biomarkers =
  mri_vars

df = 
  visit1 |> 
  select(
    `FXS ID`,
    `FX*`,
    all_of(biomarkers)) |> 
  filter(
    !is.na(`FX*`)) |> 
  mutate(
    across(all_of(biomarkers), ~ as.numeric(.x) - 1)
  ) |> 
  na.omit()

```

# Set up data for new version of SuStaIn
The objective here is two create to variables:

* `p_NL` -- a subject x region matrix indicating the p that a region is normal for a subject.
* `p_score` -- a subject x (region * score) matrix indicating p that a region is a given score.

Ps should be non-zero. 

```{r}

ModelScores = DataScores = 
  df[,biomarkers] |> unlist() |> unique(na.rm = TRUE) |> sort() |> as.character()
  # unique(df[,biomarkers], na.rm = TRUE) |> as.character()

control_data = df |> filter(`FX*` == "CGG < 55") |> select(all_of(biomarkers))
patient_data = df |> filter(`FX*` == "CGG >= 55") |> select(all_of(biomarkers))

# prob_correct = (control_data == 0) |> colMeans()
prob_correct = 
  control_data |> 
  summarize(
    across(
      .cols = all_of(biomarkers), 
      .fn = ~ mean(.x == DataScores[1], na.rm = TRUE))
  ) |> 
  unlist()

prob_correct[] = .95

prob_dist_dims = 
  list(
    modelscore = ModelScores,
    datascore = DataScores,
    biomarker = biomarkers)

prob_dist = array(
  NA,
  dim = sapply(length, X = prob_dist_dims),
  dimnames = prob_dist_dims)

for (biomarker in biomarkers)
{
  for (datascore in DataScores)
  {
    for (modelscore in ModelScores)
    {
      if(datascore == modelscore)
      {
        prob_dist[modelscore, datascore, biomarker] = prob_correct[biomarker] 
      } else
      {
        prob_dist[modelscore, datascore, biomarker] = 
          (1 - prob_correct[biomarker] ) / (length(DataScores) - 1)
      }
    }
  }
}

prob_nl_dist = prob_dist[1, , ]
prob_score_dist = prob_dist[-1, , ]

```

```{r}

prob_score_dims = 
  list(
    ID = df |> filter(`FX*` == "CGG >= 55") |> pull(`FXS ID`),
    Biomarker = biomarkers,
    model = ModelScores
  )

prob_score0 = array(
  NA,
  dim = prob_score_dims |> sapply(length),
  dimnames = prob_score_dims
)

for (biomarker in biomarkers)
{
  for (datascore in DataScores)
  {
    for (modelscore in ModelScores)
    {
      prob_score0[patient_data[[biomarker]] == datascore, biomarker, modelscore] = 
        prob_dist[modelscore, datascore, biomarker]
    }
    
  }
}

prob_nl = prob_score0[,,1]
prob_score = prob_score0[,,-1, drop = FALSE]

```

```{r "score_vals"}

score_vals = matrix(
  ModelScores[-1] |> as.numeric(),
  byrow = TRUE,
  nrow = length(biomarkers),
  ncol = length(ModelScores) - 1)



```

# Run SuStaIn model

```{r "set run parameters"}

N_startpoints = 10L
N_S_max = 3L
N_iterations_MCMC = 1e5L
dataset_name = 'sample_data'
output_folder = "output.FXTAS" |> fs::dir_create()
print(output_folder)
```

```{python "initialize-model"}

import pandas
import numpy as np
import numpy.matlib
import matplotlib.pyplot as plt
import seaborn as sns
import sys
import scipy
from glob import glob
import pySuStaIn
import os

os.getcwd()

sustain_input = pySuStaIn.OrdinalSustain(r.prob_nl,
r.prob_score,
r.score_vals,
r.SuStaInLabels,
r.N_startpoints,
r.N_S_max, 
r.N_iterations_MCMC, 
r.output_folder, 
r.dataset_name, 
False,
1,
True)

```

```{python "fit-model", fig.width = 8, fig.height = 9}

[samples_sequence, samples_f, 
ml_subtype, prob_ml_subtype, ml_stage, 
prob_ml_stage, prob_subtype_stage, 
samples_likelihood, 
fig0] = sustain_input.run_sustain_algorithm(plot=True)

```


```{r}

likelihoods = as_tibble(py$samples_likelihood) |> 
  setNames(paste(1:ncol(likelihoods), "subtype(s)")) |> 
  mutate(Iteration = row_number())

alpha = 0.5
ggplot() +
  geom_line(
    alpha = alpha,
    aes(
      x = 1:nrow(likelihoods), 
      y = likelihoods[[1]],
      col = "One subtype"
    )
  ) +
  geom_line(
    alpha = alpha,
    aes(
      x = 1:nrow(likelihoods), 
      y = likelihoods[[2]],
      col = "Two subtypes"
    )
  ) +
  geom_line(
    alpha = alpha,
    aes(
      x = 1:nrow(likelihoods), 
      y = likelihoods[[3]],
      col = "Three subtypes"
    )
  )



```

```{r}

likelihoods2 = 
  likelihoods  |> 
  tidyr::pivot_longer(cols = paste(1:ncol(py$samples_likelihood), "subtype(s)"))

likelihoods2 |> 
  ggplot(
    aes(
      x = Iteration, 
      y = value,
      col = name
    )) +
  geom_line(alpha = alpha) +
  facet_wrap(~name) +
  ylab("log-likelihood")

```


```{python}
fig0.show()
```

### trying `plt.show()`

```{python, fig.width = 8, fig.height = 9}
plt.show()
```


### attempting to call the plot function to get the heatmaps

```{python}
n_samples = r.prob_nl.shape[0]

```

```{python, include = TRUE, fig.width = 8, fig.height = 9}
# plt.show()

```

```{python}

# plot_positional_var()

```


```{python}

figs, ax = sustain_input.plot_positional_var(
samples_sequence=samples_sequence,
samples_f=samples_f,
n_samples= n_samples,
score_vals = r.score_vals,
biomarker_labels=r.biomarkers,
save_path=r.output_folder)
```



```{python, "show figs"}
for fig in figs:
fig.show()

# py$samples_sequence

```



