---
title: "PAL Total errors"
knitr:
  opts_chunk:
    echo: false
    message: false
    warning: false
---

```{r setup}
#| include: false
library(fxtas)
library(dplyr)
```

@fig-PAL-trax-gp34-males shows "PAL Total errors" by age at visit and CGG repeats, for male participants in GP, GP4, and Trax. The black horizontal line indicates the currently used cutoff (PAL = 13). The colored lines connect multiple visits from the same participant. Hovering over the individual observations will produce pop-up windows specifying x and y coordinates and ID#s. You can zoom and pan by hovering over the axes or using the buttons that appear on hover in the top right.

@tbl-pal provides summary statistics for "PAL Total errors" by CGG level.

```{r}
#| label: fig-PAL-trax-gp34-males
#| fig-cap: "PAL Total errors, by age at visit and CGG repeats level, for males in GP, GP4, and Trax"
# males_gp34_trax_v1 |> 
#   select(Study, `PAL Total errors (adjusted)`) |> 
#   table(useNA = 'ifany')

library(ggplot2)
pd <- position_dodge(2)
plot1 = 
  males_gp34_trax |> 
  mutate(
    Study2 = if_else(
      Study %in% c("GP", "GP4"),
      "GP3+GP4",
      "Trax"
    )
  ) |> 
  arrange(`FXS ID`, `Visit Date`) |> 
  
  filter(
    # `FXS ID` == "500011-138", # most frequently observed
    !is.na(`PAL Total errors (adjusted)`),
    !is.na(`FX*`)
  ) |> 
  ggplot(
    aes(
      col = `FXS ID`,
      # group = `FXS ID`,
      x = `Age at visit`, 
      y = `PAL Total errors (adjusted)`)) +
  geom_point(alpha = .5, position = pd) +
  geom_line(alpha = .5, position = pd) +
  facet_wrap(~`FX3*`) +
  geom_hline(yintercept = 13) +
  theme(legend.position = "none") +
  labs(colour = "")

# print(plot1)
library(plotly)
ggplotly(plot1)
```

```{r}
#| label: tbl-pal
#| tbl-cap: "PAL Total errors by CGG repeats level, for males in GP, GP4, and Trax"

library(table1)
males_gp34_trax |> 
  filter(!is.na(`FX*`)) |>  
  mutate(
    Study2 = if_else(
      Study %in% c("GP", "GP4"),
      "GP3+GP4",
      "Trax"
    )
  ) |> 
  table1(
    ~ `PAL Total errors (adjusted)` | `FX3*`,
    # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
    data = _)

```

