---
title: "potential variables list"
format: 
  html:
    embed-resources: true
editor: visual
---


```{=html}
<style>
.quarto-figure-center > figure {
  text-align: center;
}
</style>
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  # fig.width = 7, 
  # fig.height = 10,
  include = TRUE
)
```

```{r setup}
#| message: false
devtools::load_all()
library(tidyverse)
library(reticulate)
library(pander)
```

```{r}

biomarker_group_list = 
  all_potential_biomarkers()

biomarker_group_colors = 
  tibble(
    biomarker_group = names(biomarker_group_list),
    # group_color = rainbow()
    group_color = 
      length(biomarker_group_list) |> 
      # RColorBrewer::brewer.pal(n = _, name = "Paired")
      colorspace::qualitative_hcl(n = _, "Dark2")
  )

biomarker_groups = 
  biomarker_group_list |> 
  stack() |> 
  as_tibble() |> 
  rename(
    biomarker  = values,
    biomarker_group = ind) |> 
  left_join(
    biomarker_group_colors,
    by = "biomarker_group"
  ) |> 
  mutate(
    biomarker_group = 
      biomarker_group |> 
      factor(levels = names(biomarker_group_list)))

```

@tbl-biomarker-list lists the biomarkers that are potentially usable for analysis by the Ordinal SuStaIn method.

Notes: 

- "Any Autoimmune" includes:
  - "Lupus",
  - "Rheumatoid arthritis",
  - "Multiple Sclerosis: Workup",
  - "ANA positive",
  - "Sjogrens Syndrome",
  - "Raynauds Syndrome",
  - "Pulmonary Fibrosis"
  
- "Any tremor (excluding Head Tremor)" includes:
  - "Intention tremor",
  - "Resting tremor",
  - "Postural tremor",
  - "Intermittent tremor"
  
- "Any cancer" includes:
  - "Thyroid Cancer",
  - "Skin Cancer",
  - "Melanoma",
  - "Prostate Cancer",
  - "Other Cancer"

- an asterisk ("*") after the biomarker name indicates a numerical variable that has been categorized.

```{r}
#| tbl-cap: "Potential biomarkers for Ordinal SuStaIn analysis"
#| label: "tbl-biomarker-list"
table_out =
  biomarker_groups |> 
  select(category = biomarker_group, biomarker) |> 
  slice_head(by = biomarker)

table_out |> 
  pander()

```
