---
title: "Event sequences in Fragile X-associated tremor/ataxia syndrome"
format: 
  html: 
    author: 
      - name: Ezra Morrison
        email: demorrison@ucdavis.edu
      - Matt Ponzini
      - Kyoungmi Kim
      - (more to be added)
    toc: true
    embed-resources: true
    mainfont: serif
    theme:
       light: 
         - cosmo
         - custom.scss
       dark: 
         - darkly
         - custom.scss
    cap-location: top
    css: custom.scss
    number-sections: true
    number-depth: 3
    html-math-method: mathjax
    fig-cap-location: top
    code-fold: true
    echo: false
    date: last-modified
    date-format: "[Last modified:] YYYY-MM-DD: H:mm:ss (A)"
    bibliography: references.bib

knitr:
  opts_chunk: 
    collapse: true
    comment: "#>" 
    R.options:  
      dev: svg

project:
  execute-dir: project
---

```{=html}
<style>
.quarto-figure-center > figure {
text-align: center;
}
</style>
```

# Methods

Here, we fit the Ordinal SuStaIn model [@young2021ordinal] on three datasets:

- GP + GP4, males only, first visit only
- Trax + GP + GP4, males only, first visit only
- Trax (already all males), first visit only

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # fig.width = 7, 
  # fig.height = 10,
  include = TRUE
)
```

```{r setup}
#| message: false
devtools::load_all()
# library(fxtas)
library(tidyverse)
library(reticulate)
use_condaenv("fxtas39")
# py_config()
library(pander)
```

```{r "build-figs"}
fig_male = extract_figs_from_pickle(
  n_s = 1,
  dataset_name = "males",
  output_folder = "output/output.longest")
```

```{r}
fig_male_trax_gp34 = extract_figs_from_pickle(
  n_s = 1,
  dataset_name = "sample_data",
  output_folder = "output/trax_gp34_males_v1")

fig_trax_only = extract_figs_from_pickle(
  n_s = 1,
  dataset_name = "trax_visit1",
  output_folder = "output/trax_visit1")
```

# Results

@fig-pvd1 shows the event sequences that we estimated from each dataset. Note that clinical events which never occur in a dataset cannot be placed into the event sequence, so the set of modeled events varies by data subset. Additionally, since we only used the earliest visit for each individual across all the datasets in each analysis, some of the events included in the GP+GP4 1st visits dataset were not observed in the GP+GP4+Trax 1st visits dataset.

```{r "extract-event-lists"}
gp34_events = fig_male |> attr("biomarker_event_names")
trax_events = fig_trax_only |> attr("biomarker_event_names")
trax_gp34_events = fig_male_trax_gp34 |> attr("biomarker_event_names")
```

```{r}
missing1 = setdiff(gp34_events, trax_events)
missing2 = setdiff(gp34_events, trax_gp34_events)
```

Specifically,  compared to the GP+GP4 dataset, the Trax+GP+GP4 dataset included `r length(missing2)` fewer event types:

```{r}
print(missing2 |> sort())
```

Compared to the GP+GP4 1st visits dataset, the Trax data included `r length(missing1)` fewer event types:

```{r}
print(missing1 |> sort())
```



```{r}
#| include: false
library(VennDiagram)
temp = venn.diagram(
  x = list(gp34_events, trax_events, trax_gp34_events),
  category.names = c("GP+GP4" , "Trax" , "GP+GP4+Trax"),
  filename = "output/venn_dagram.png",
  output=TRUE
)
```

```{r}
#| label: "fig-pvd1"
#| column: screen
#| fig-height: 15
#| fig-align: center
#| fig-cap: "Estimated event sequence models - males only: gp34, trax, and gp34+trax"
#| fig-cap-location: top
#| layout-ncol: 3
#| fig-subcap:
#|   - "GP+GP4 males only"
#|   - "Trax+GP+GP4, males only" 
#|   - "Trax only"

fig_male = extract_figs_from_pickle(
  n_s = 1,
  dataset_name = "males",
  output_folder = "output/output.longest") |> 
  print()

fig_male_trax_gp34 = extract_figs_from_pickle(
  n_s = 1,
  dataset_name = "sample_data",
  output_folder = "output/trax_gp34_males_v1") |> 
  print()

fig_trax_only = extract_figs_from_pickle(
  n_s = 1,
  dataset_name = "trax_visit1",
  output_folder = "output/trax_visit1") |> 
  print()

```

# References {.unnumbered}

::: {#refs}
:::
