---
title: "Questions about FXTAS Sequence Data"
subtitle: "WAIS"
format:
  html:
    toc: false
    toc-depth: 4
    toc-float: true
    number-sections: false
    embed-resources: true
---

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7.5,
  fig.height = 4
)

# update 'asis' chunk to allow inline code
knitr::knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knitr::knit_child(text = options$code)
})
```

```{r setup}
#| message: false
library(fxtas)
devtools::load_all()
library(pander)
library(table1)
library(dplyr)
library(magrittr)
NA.label1 = "Missing"
column_var = "Recruited in study phase"
column_var = "FX**"
# load_all()

phase <- ifelse(column_var == "Recruited in study phase", TRUE, FALSE)
cgg <- ifelse(column_var == "FX**", TRUE, FALSE)
```

We combined "Verbal: IQ Score" (`wais_verb_iq`) from the GP study, which used WAIS3, and "Verbal Comprehension: Composite Score (VCI)" (`wais4_verbcomp_cs`) from the GP4, which used WAIS4. Since these scores are already standardized, we treated them as interchangeable. 

Similarly, we combined  "Full Scale: IQ Score" (`wais_fullscale_iq`) from the GP study and 
"Full Scale: Composite Score (FSIQ)" (`wais4_fullscale_cs`) from the GP4 study, treating them as interchangeable.

```{r}

wais_vars0 =
  c("Verbal Comprehension: Composite Score (VCI)",
  
  "Perceptual Reasoning: Composite Score (PRI)",
  "Working Memory: Composite Score (WMI)", 
  "Processing Speed: Composite Score (PSI)",
  
  # "Full Scale: IQ Score",
  "Full Scale: Composite Score (FSIQ)")
vars = wais_vars0 |> sapply(F = function(x) grep(value =  TRUE, fixed = TRUE, pattern = x, x = names(visit1))) |> as.vector()

table1(
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min, Max]"),
  data = visit1)

```


```{r}
missing_wais = 
  gp34 |> 
  filter(
    if_any(all_of(wais_vars0), is.na)
  ) |> 
  select(
    `FXS ID`,
    `Event Name`,
    all_of(wais_vars0)
  )
readr::write_csv(missing_wais, "output/incomplete_wais.csv")

```

