@suppfig-pvd-by-cgg shows the estimated sequences 
stratified by CGG repeat level.
We did not find statistically significant evidence of a difference in event sequences
between CGG < 100 and CGG ≥ 100 (p = `r pval_cgg_gp34_v1`),
but some potential differences were noted
(@suppfig-pvd-by-cgg and @suppfig-diff-by-cgg).
SWM between errors occurred prior to Stage 1
in participants with CGG repeats <100,
whereas it occurred between stage 4 and 5 in those with CGG repeats ≥ 100
(@suppfig-diff-by-cgg).
Postural, resting, head, and intermittent tremor occurred earlier
in those with CGG repeats <100 compared to those with CGG repeats ≥100.
Several psychiatric disorders (as measured by SCID) occurred
later in the event sequence in participants with CGG repeats <100
compared to those with CGG repeats ≥100.

The finding of only minimal differences 
between 55-99 vs 100-199 CGG repeats is surprising, 
since several studies have found that the higher the CGG repeats,
the earlier the onset and the faster the progression of FXTAS 
[@greco2006neuropathology; @leehey2008fmr1; @tassone2023insight]. 
Perhaps the cut off of 100 is too high for this distinction to be made.

These differences were more evident among males than females
(Supplementary material, [Comparing CGG levels stratified by sex](#sec-cgg-strat-sex)).
When we fitted models for subgroup analyses by sex and CGG repeats level,
the models for males showed the same patterns described above,
except for head tremor
(we found no difference in head tremor onset between CGG levels).
@suppfig-pvd-by-cgg-male
shows the corresponding stratified models and
compares the estimates of the sequences for these two subgroups.


We did not find statistically significant evidence of
a difference between CGG <100 and CGG ≥100 among females
(p = `r pval_female_cgg_gp34_v1`).
@suppfig-pvd-by-cgg-female
shows the estimated stratified models
and compares the estimates.
Females had later onset of SWM Between Errors for CGG repeats ≥100
and later onset of intermittent tremors for CGG repeats <100,
but did not show evidence of the other patterns observed for males and
for the models without sex-stratification.
Limited sample size in females for subgroup analyses 
was likely an issue;
the available data included only `r n_over100_Female` females with CGG ≥100.

{{< pagebreak >}}

::: {#suppfig-pvd-by-cgg}

```{r}
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width

list(
    fig_under100,
    fig_over100
  ) |> 
plot_compact_pvd(
  facet_label_prefix = c("CGG &lt; 100 <br>", "CGG &ge; 100 <br>"),
  y_text_size = y_text_size
)


```

`r glue::glue("**Event sequences stratified by CGG repeats (<100 vs 100+).** {compact_fig_caption}")`
:::

::: {#suppfig-diff-by-cgg}

```{r}
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width


list(
  "CGG &lt; 100" = fig_under100,
  "CGG &ge; 100" = fig_over100) |>
pvd_lineplot(y_text_size = y_text_size)
```

**Event sequences stratified by CGG repeats (<100 vs 100+).** 
Differences in event sequences between CGG repeat sizes.
`r compact_fig_cap2`

:::


{{< pagebreak >}}
