The following supplementary tables 
[-@supptbl-tremors] - [-@supptbl-autoimmune] 
summarize the distributions of the `r outcomes_name`s
that we analyzed in this paper, stratified by sex and CGG level.

```{r}
#| label: exploration-setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7.5,
  fig.height = 4
)
```

```{r}
#| label: setup-demos-unstratified
#| message: false
library(fxtas)
library(pander)
library(table1)
library(dplyr)
library(magrittr)

# dealing with missing values:
na_label_1 <- "Missing"
show_missingness_reasons <- FALSE
footnote1 <- 'Codes `777`, `888`, and `999` are counted as "Missing" (NA).'
```


::: {#supptbl-tremors}

```{r}
#| label: tab-tremors
vars <- c(
  "Intention tremor",
  "Resting tremor",
  "Postural tremor",
  "Intermittent tremor"
)
table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  render.continuous = c(
    . = "N",
    . = "Mean (SD)",
    . = "Median [Min, Max]"
  ),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

**Tremors by CGG level and sex**

:::

::: {#supptbl-head-tremors}

```{r}
vars <- c(
  "Head tremor"
  # "Head Tremor: Age of onset"
)

footnote_ht <- paste(
  "Head tremor was not collected in GP3 or Trax Phase 1.",
  "Summaries provided are for GP4 + Trax Phase 2 only."
)
v1_usable |>
    # dplyr::filter(!(Study %in% c("Trax Phase 1", "GP3"))) |> 
table1(
  # footnote = footnote_ht,
  na.is.category = FALSE,
  overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = _
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

**Head tremors**

:::

::: {#supptbl-ataxia-by-cgg}

```{r}
vars <- c(
  "Ataxia",
  # "Ataxia: severity",
  "Ataxia: severity*"
  # "Ataxia: Age of onset",
  # "Ataxia: Age of onset: missingness reasons"
)

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

Ataxia by CGG status

:::

::: {#supptbl-parkinsons-by-cgg}

```{r}
vars <- grep("Parkinson", value = TRUE, names(trax_gp34_all))

if (!show_missingness_reasons) {
  vars <- vars |> remove_missingness_reason_vars()
}

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

Parkinson's disease by CGG status

:::

::: {#supptbl-fxtas-stage-by-cgg}

```{r}
vars <- c(
  "FXTAS Stage",
  if (show_missingness_reasons) "FXTAS Stage: missingness reasons"
)

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()

```

FXTAS Stage by CGG level

:::

::: {#supptbl-bds2}

```{r}
vars <- c("BDS-2 Total Score", "BDS-2 Total Score*")

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  # footnote = footnote1,
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

Behavior Dyscontrol Scale - Second Edition (BDS-2)

:::

::: {#supptbl-mmse}

```{r}
# vars = grep(
#   "MMSE total score",
#   value = TRUE,
#   names(trax_gp34_all),
#   fixed = TRUE)

vars <- c(
  "MMSE total score",
  # "MMSE total score: missingness reasons",
  "MMSE total score*"
)

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  # footnote = footnote1,
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

Mini-Mental State Examination (MMSE)

:::

::: {#supptbl-scid}

```{r}
vars <- scid_vars_for_table1()

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

**Structured Clinical Interview for DSM Disorders (SCID)**

:::

::: {#supptbl-mri}

```{r}
vars <- c(
  # "Cerebral Atrophy",
  # "Cerebellar Atrophy",
  # "Cerebral WM Hyperintensity",
  # "Cerebellar WM Hyperintensity",
  # "MCP-WM Hyperintensity",
  # "Pons-WM Hyperintensity",
  # "Sub-Insular WM Hyperintensity",
  # "Periventricular WM Hyperintensity",
  "MRI: Cerebellar",
  "MRI: Cerebral",
  "Splenium (CC)-WM Hyperintensity",
  "Genu (CC)-WM Hyperintensity",
  "Corpus Callosum-Thickness"
)

table1(
  na.is.category = FALSE,
  overall = FALSE,
  NA.label = na_label_1,
  droplevels = FALSE,
  x = stratified_formula(vars, column_var),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

**MRI variables by Sex and CGG level**

:::

::: {#supptbl-cantab}

```{r}
# table1(na.is.category = FALSE, overall = FALSE,
#   NA.label = na_label_1,
#   stratified_formula("Drugs used", column_var),
#   data = v1_usable)


cantab_vars <- c(
  "SWM Between errors",
  "SWM Between errors*",
  # "SST Median correct RT on GO trials",
  # "RVP A signal detection",
  # "RVP A signal detection*",
  # "OTS Problems solved on first choice",
  "PAL Total errors (adjusted)",
  "PAL Total errors (adjusted)*",
  "RTI Five-choice movement time",
  "RTI Five-choice movement time*"
)
# >
# sapply(
#   F = function(x)
#     grep(
#       value =  TRUE,
#       fixed = TRUE,
#       pattern = x,
#       x = names(v1_usable)
#     )
# ) |>
#   as.vector()


  table1(
    na.is.category = FALSE, overall = FALSE,
    NA.label = "Missing (empty in RedCap)",
    stratified_formula(cantab_vars, column_var),
    # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
    data = v1_usable
  ) |> 
    table1::t1flex() |>
    shared_flextable_settings() |>
    flextable::width(width = 2) |>
    flextable::width(j = 2:7, width = 1.25)
```

**Cambridge Neuropsychological Test Automated Battery (CANTAB)**

:::

We used categorization cutoffs taken from Talebi et al 2020 [@talebi2020cambridge].


::: {#supptbl-autoimmune}

```{r}
vars <- c(
  "Hypothyroid",
  "Hyperthyroid",
  # "Thyroid problems",
  # "Lupus",
  # "Rheumatoid arthritis",
  # "Multiple Sclerosis: Workup",
  # "ANA positive",
  # "Sjogrens Syndrome",
  # "Raynauds Syndrome",
  # "Pulmonary Fibrosis",
  "any autoimmune disorder"
)

table1(
  na.is.category = FALSE,
  overall = FALSE,
  NA.label = na_label_1,
  stratified_formula(vars, column_var),
  data = v1_usable
) |> 
  table1::t1flex() |>
  shared_flextable_settings()
```

**Thyroid and autoimmune diseases**

:::


