---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
#| label: exploration-setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7.5,
  fig.height = 4
)
```

```{r}
#| label: setup-demos-unstratified
#| message: false
library(fxtas)
# devtools::load_all()
library(pander)
library(table1)
library(dplyr)
library(magrittr)

# dealing with misssing values:
NA.label1 = "Missing"
show_missingness_reasons = FALSE
footnote1 = 'Codes `777`, `888`, and `999` are counted as "Missing" (NA).'

```


### Tremors {.appendix}

```{r}
#| label: tbl-tremors
#| tbl-cap: "Tremors"
vars = c(
  
  "Intention tremor",
  "Resting tremor",
  "Postural tremor",
  "Intermittent tremor",
  "Any tremor (excluding Head Tremor)"
  # "Tremor: Age of onset"
  # if(show_missingness_reasons) "Tremor: Age of onset: missingness reasons"
)
table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       render.continuous=c(
         .="N", 
         .="Mean (SD)", 
         .="Median [Min, Max]"),
       data = v1_usable)

```

<br>

```{r}
#| tbl-cap: "Age of onset for those with 'Some tremors recorded'."
#| label: "tbl-tremor-ao"
#| 
vars = c(
         "Tremor: Age of onset",
         # "Tremor: Age of onset: missingness reasons",
         NULL
         )


table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = v1_usable |> 
         filter(`Any tremor (excluding Head Tremor)` == "Some tremors recorded"))
```

<br>

### Head tremors  {.appendix}

```{r}
#| tbl-cap: "Head tremors"
#| label: tbl-head-tremors
vars = c(
  "Head tremor"
  # "Head Tremor: Age of onset"
  
)

footnote_HT <- paste(
  "Head tremor was not collected in GP3 or Trax Phase 1.",
  "Summaries provided are for GP4 + Trax Phase 2 only."
)

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = v1_usable |> 
    filter(!(Study %in% c("Trax Phase 1", "GP3"))),
  footnote = footnote_HT)

```

<br>

```{r}
#| tbl-cap: "Age of onset for those with a head tremor recorded."
#| label: "tbl-head-tremor-AO"
#| eval: false # converting to docx runs into a weird `flextable()` edge case because two subcolumn headers are identical
vars = c(
  "Head Tremor: Age of onset",
  if(show_missingness_reasons) "Head Tremor: Age of onset: missingness reasons"
  
)

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = v1_usable |> 
    filter(
      !(Study %in% c("Trax Phase 1", "GP3")),
      `Head tremor` == "Yes"))

```

### Ataxia  {.appendix}

```{r}
#| label: tbl-ataxia-by-cgg
#| tbl-cap: "Ataxia by CGG status"
vars = c(
  "Ataxia",
  # "Ataxia: severity",
  "Ataxia: severity*"
  # "Ataxia: Age of onset",
  # "Ataxia: Age of onset: missingness reasons"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = v1_usable)

```

<br>

```{r}
#| label: tbl-ataxia-AO-by-cgg
#| tbl-cap: "Ataxia age of onset by CGG level"
vars = c(
  # "Ataxia",
  # "Ataxia: severity",
  # "Ataxia: severity*"
  "Ataxia: Age of onset"
  # "Ataxia: Age of onset: missingness reasons"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = v1_usable |> filter(Ataxia == "Yes"))
```

<br>

### Parkinsons  {.appendix}

```{r}
#| label: tbl-parkinsons-by-cgg
#| tbl-cap: "Parkinsons by CGG status"
vars = grep("Parkinson", value = TRUE, names(trax_gp34_all))

if(!show_missingness_reasons) 
{
  vars = vars |> remove_missingness_reason_vars()
}

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       data = v1_usable)

```

### FXTAS Stage {.appendix}

```{r}
#| label: tbl-fxtas-stage-by-cgg
#| tbl-cap: "FXTAS Stage by CGG level"
vars = c(
  "FXTAS Stage (0-5)*",
  if(show_missingness_reasons) "FXTAS Stage (0-5)*: missingness reasons"
)

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       footnote = "For categorical analysis, we merged 2.5 into 2, 3.5 into 4, and 4.5 into 5.",
       data = v1_usable)

# trax_gp34_all |> 
#   filter(`FXTAS Stage (0-5)` %in% c(NA, "No Response (999)", "Missing (empty in RedCap)")) |> 
#   select(`FXS ID`, `Event Name`, `Visit Date`) |> 
#   arrange() |> 
#   readr::write_csv("inst/extdata/missing_FXTAS_stage.csv")

```

### Behavior Dimensions Scale - Second Edition (BDS-2) {.appendix}

```{r}
#| label: tbl-bds2
#| tbl-cap: "Behavior Dimensions Scale - Second Edition (BDS-2)"
vars = c("BDS-2 Total Score", "BDS-2 Total Score*")


table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       footnote = footnote1,
       data = v1_usable)

```

### Mini-Mental State Examination (MMSE)  {.appendix}

```{r}
#| label: tbl-mmse
#| tbl-cap: "Mini-Mental State Examination (MMSE)"
# vars = grep(
#   "MMSE Total Score", 
#   value = TRUE, 
#   names(trax_gp34_all), 
#   fixed = TRUE)

vars = c(
  "MMSE Total Score", 
  # "MMSE Total Score: missingness reasons", 
  "MMSE Total Score*")

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = NA.label1,
       stratified_formula(vars, column_var),
       footnote = footnote1,
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = v1_usable)

```

### Structured Clinical Interview for DSM Disorders (SCID)  {.appendix}

```{r}
#| label: tbl-scid
#| tbl-cap: "Structured Clinical Interview for DSM Disorders (SCID)"
vars = scid_vars_for_table1()

table1(
  na.is.category = FALSE, overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
  data = v1_usable)

```



### MRI  {.appendix}

```{r}
#| label: tab-mri
#| tbl-cap: "MRI variables by Sex and CGG level"
vars = c(
  # "Cerebral Atrophy",
  # "Cerebellar Atrophy",
  # "Cerebral WM Hyperintensity",
  # "Cerebellar WM Hyperintensity",
  # "MCP-WM Hyperintensity",
  # "Pons-WM Hyperintensity",
  # "Sub-Insular WM Hyperintensity",
  # "Periventricular WM Hyperintensity",
  "MRI: Cerebellar",
  "MRI: Cerebral",
  "Splenium (CC)-WM Hyperintensity",
  "Genu (CC)-WM Hyperintensity",
  "Corpus Callosum-Thickness"
)

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  data = v1_usable)

```

### Wechsler Adult Intelligence Scale (WAIS)  {.appendix}

```{r}
#| label: tbl-wais
#| tbl-cap: "Wechsler Adult Intelligence Scale (WAIS)"
vars = c(
  # "Verbal: IQ Score",
  "Verbal Comprehension: Composite Score (VCI)",
  
  "Perceptual Reasoning: Composite Score (PRI)",
  "Working Memory: Composite Score (WMI)", 
  "Processing Speed: Composite Score (PSI)",
  
  # "Full Scale: IQ Score",
  "Full Scale: Composite Score (FSIQ)"
)

if(show_missingness_reasons)
{
  vars = vars |> 
    sapply(
      F = function(x) 
        grep(
          value =  TRUE, 
          fixed = TRUE, 
          pattern = x, 
          x = names(v1_usable))) |> as.vector()
}

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min, Max]"),
  data = v1_usable)

```

### Cambridge Neuropsychological Test Automated Battery (CANTAB)  {.appendix}

We used categorization cutoffs taken from [@talebi2020cambridge].

```{r}
#| label: tbl-cantab
#| tbl-cap: "Cambridge Neuropsychological Test Automated Battery (CANTAB)"
#| 
# table1(na.is.category = FALSE, overall = FALSE,
#   NA.label = NA.label1,
#   stratified_formula("Drugs used", column_var),
#   data = v1_usable)


cantab_vars = c(
  "SWM Between errors",
  "SWM Between errors*",
  "SST Median correct RT on GO trials",
  "RVP A signal detection",
  "RVP A signal detection*",
  "OTS Problems solved on first choice",
  "PAL Total errors (adjusted)",
  "PAL Total errors (adjusted)*",
  "RTI Five-choice movement time",
  "RTI Five-choice movement time*") 
# |> 
#   sapply(F = function(x) grep(value =  TRUE, fixed = TRUE, pattern = x, x = names(v1_usable))) |> 
# as.vector()

table1(na.is.category = FALSE, overall = FALSE,
       NA.label = "Missing (empty in RedCap)",
       stratified_formula(cantab_vars, column_var),
       # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
       data = v1_usable)
```

### Thyroid and autoimmune diseases  {.appendix}

```{r}
#| label: tbl-autoimmune
#| tbl-cap: "Thyroid and autoimmune diseases"
vars = c(
  "Hypothyroid",
  "Hyperthyroid",
  # "Thyroid problems",
  # "Lupus",
  # "Rheumatoid arthritis",
  # "Multiple Sclerosis: Workup",
  # "ANA positive",
  # "Sjogrens Syndrome",
  # "Raynauds Syndrome",
  # "Pulmonary Fibrosis",
  "Any Autoimmune"
)

table1(
  na.is.category = FALSE, 
  overall = FALSE,
  NA.label = NA.label1,
  stratified_formula(vars, column_var),
  data = v1_usable)
```
