Here we compared the longitudinal models to the cross-sectional models (estimated using each participant's earliest visit):

- Males only (@fig-males):
  * first visits only (@fig-males)
  * all visits (@fig-males)

- Females only (@fig-females):
  * first visits only (@fig-females)
  * all visits (@fig-females)

We also rearranged these results by subset type:

- first visits only (@fig-first-only)
- all visits (@fig-all-visits)

Note that clinical events which never occur in a dataset cannot be placed into the event sequence, 
so the set of modeled events varies by data subset.

```{r}
#| label: extract-figs-longitud
fig_females_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "females",
  output_folder = "output/gp34_longitudinal")

fig_females_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "females",
  output_folder = "output/gp34_multivisit_only")


fig_males_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "males",
  output_folder = "output/gp34_longitudinal")

fig_males_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "males",
  output_folder = "output/gp34_multivisit_only")


fig_both_all = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "gp34_longitudinal",
  output_folder = "output/gp34_longitudinal")

fig_both_mv_only = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "gp34_multivisit_only.RData",
  dataset_name = "gp34_multivisit_only",
  output_folder = "output/gp34_multivisit_only")

```

### Males {.appendix}

@fig-males shows estimated event sequences for males, depending on which subset of the data we use to fit the model.

```{r}
#| label: fig-males
#| column: page
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| fig-cap: !expr glue::glue("Estimated event sequences for males, by data subset. {compact_fig_caption}")
#| fig-cap-location: top

list(
    "Males, First Visit Only" = fig_males_first,
    "Males, All Visits" = fig_males_all
  ) |> 
plot_compact_pvd(y_text_size = 11)

```

{{< pagebreak >}}

```{r}
#| label: "fig-males-change"
#| column: page
#| fig-height: !expr pvd_line_height
#| fig-width: !expr pvd_line_width
#| fig-cap: "Change in estimated event sequences for males by data subset"
#| fig-cap-location: top

list(
    "Males, First Visit Only" = fig_males_first,
    "Males, All Visits" = fig_males_all
  ) |> 
pvd_lineplot(y_text_size = 11)
```

### Females {.appendix}

```{r}
#| label: fig-females
#| fig-cap: !expr glue::glue("Estimated event sequences for females, by data subset. {compact_fig_caption}")
#| column: page
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| fig-cap-location: top

list(
    "Females, First Visit Only" = fig_females_first,
    "Females, All Visits" = fig_females_all
  ) |> 
plot_compact_pvd(y_text_size = 11)
```

{{< pagebreak >}}

```{r}
#| label: "fig-females-change"
#| column: page
#| fig-height: !expr pvd_line_height
#| fig-width: !expr pvd_line_width
#| fig-cap: "Change in estimated event sequences for females by data subset"
#| fig-cap-location: top

list(
    "Females, First Visit Only" = fig_females_first,
    "Females, All Visits" = fig_females_all
  ) |> 
pvd_lineplot(y_text_size = 11)
```

### All patients, all visits  {.appendix}

@fig-all-visits shows the event sequences that we estimated from each data subset.
Parkinsonian features mostly appear after FXTAS stage 4 for females, 
whereas they occur before stage 4 for males.
Similarly, clinically-elevated SCID values appear before FXTAS stage 1 for females, 
whereas they occur after stage 3 for males.

{{< pagebreak >}}

```{r}
#| label: "fig-all-visits"
#| column: page
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| fig-cap: !expr glue::glue("Estimated event sequences - all visits. {compact_fig_caption}")
#| fig-cap-location: top

list(
    "Males" = fig_males_all,
    "Females" = fig_females_all
  ) |> 
plot_compact_pvd(y_text_size = 11)
```

{{< pagebreak >}}

```{r}
#| label: "fig-all-visits-change"
#| column: page
#| fig-height: !expr pvd_line_height
#| fig-width: !expr pvd_line_width
#| fig-cap: "Change in estimated event sequences - all visits"
#| fig-cap-location: top

list(
    "Males" = fig_males_all,
    "Females" = fig_females_all
  ) |> 
pvd_lineplot(y_text_size = 11)
```

{{< pagebreak >}}

```{r}
#| label: fig-stage-by-age-longitudinal
#| fig-cap: "Progression stage by age, estimated using longitudinal data"
#| include: true

patient_data1 = 
  trax_gp34_all |>
  filter(
    CGG < 200, 
    CGG >= 55)

results_both_all = extract_results_from_pickle(
  n_s = 1,
  rda_filename = "gp34_longitudinal.RData",
  dataset_name = "gp34_longitudinal",
  output_folder = "output/gp34_longitudinal")

stages = 
  results_both_all$subtype_and_stage_table |> 
  mutate(
    age = patient_data1$`Age at visit`,
    id = patient_data1$`FXS ID`
  ) |> 
  arrange(id, age)

library(ggplot2)
stages |> graph_stage_by_age()

```
