---
title: "Event sequences in Fragile X-associated tremor/ataxia syndrome"
knitr:
  opts_chunk: 
    collapse: true
    comment: "#>" 
project: 
  execute-dir: project
filters:
  - ../docx-landscape.lua
---

```{=html}
<style>
.quarto-figure-center > figure {
text-align: center;
}
</style>
```

# Introduction

The fragile X-associated tremor/ataxia syndrome (FXTAS) is a relatively newly discovered neurodegenerative disorder caused by the premutation (55 to 200 CGG repeats) in the FMR1 gene [@hagerman2001intention]. 
The premutation is associated with elevation of the FMR1 mRNA, and this leads to RNA toxicity including oxidative stress, mitochondrial dysfunction, calcium dysregulation and sequestration of proteins important to neuronal function.
Proteins and neurofilaments sequester with the excess FMR1 mRNA, leading to the formation of eosinophilic, tau negative intranuclear inclusions in neurons and astrocytes throughout the CNS and even in the peripheral nervous system [@greco2006neuropathology; @hunter2012capturing]. 
FXTAS usually begins in the 60s, and as premutation carriers age, the prevalence of FXTAS increases from approximately 40% in males in their 60s to 75% in their 80s [@jacquemont2004penetrance].
Females with the premutation are also at risk for FXTAS, but the prevalence is lower; approximately 16 to 20% may develop FXTAS, but it is less severe than in the males [@tassone2023insight]. 
The higher the CGG repeat within the premutation range the earlier the onset of FXTAS and the earlier the age of death [@greco2006neuropathology].
The prevalence of the premutation is approximately 1 in 148 to 200 females in the general population and one in 290 to 855 males (@maenner2013fmr1; @hunter2014epidemiology).

A currently open question is: In what sequential time-order do FXTAS symptoms typically appear, how does that sequence vary between participants, and what baseline characteristics predict that order?
Here, we apply the Ordinal **Su**btype and **St**age **In**ference ("SuStaIn") algorithm [@young2021ordinal] to find event sequences for FXTAS patients.

```{r, include = FALSE}
#| label: setup
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # fig.width = 7, 
  # fig.height = 10,
  include = TRUE
)
```

```{r libraries}
#| message: false
# devtools::load_all()
library(fxtas)
reticulate::use_condaenv("fxtas39",required = TRUE)
library(tidyverse)
library(reticulate)
library(pander)
library(table1)
```

```{r}
#| label: "set run parameters"
#| 
fit_models = TRUE
# fit_models = FALSE
run_CV =  TRUE
# run_CV = FALSE

pvd_height = 8
size.y = 11
N_startpoints = 10L
N_S_max = 8L
N_S_max_stratified = 1L
N_CV_folds = 10L

# column_var = "Recruited in study phase"
# column_var = "FX3*"
column_var = c("Gender", "FX3*")

N_iterations_MCMC = 1e5L
dataset_name = 'sample_data'
root_dir = here::here()
setwd(root_dir)
output_folder = 
  "output/output.fixed_CV/" |> 
  fs::dir_create()
```

# Methods

## Data

```{r}
#| label: subset-data

full_data = trax_gp34_v1
n_missing_CGG = full_data$CGG |> is.na() |> sum()
n_above_200 = sum(full_data$CGG >= 200, na.rm = TRUE)
v1_usable = full_data |> filter(CGG < 200)
v1_usable_cases = v1_usable |> filter(CGG >= 55)
# note: there are 231 records in `visit1` with CGG >= 55, but 4 have CGG >= 200
# previously `nrow(v1_usable_cases)` was 221, which was based on incorrectly filtering on a version of CGG that hadn't been backfilled.

gp34_v1_controls =  v1_usable |> filter(CGG < 55)
```

### Study cohorts

The data come from the GP, GP4, and Trax studies. In the primary analyses, we analyzed one visit record per participant, using the earliest visit available across the two studies with age at visit $\ge 40$ years, resulting in `r nrow(full_data)` observations in total. 
However, `r n_missing_CGG` individuals did not have CGG values available, and `r n_above_200` had full fragile X syndrome mutation (≥200 CGG repeats), so only `r nrow(v1_usable)` individuals 
(`r nrow(v1_usable_cases)` cases and 
`r nrow(gp34_v1_controls)` controls) were included in the event sequence analysis. 
@fig-hist-CGG-gp34-v1 shows the distribution of CGG repeats among the participants included in the analysis. 
@tbl-demographics provides demographic information about the study participants. 

```{r}
#| label: "biomarker-events"
biomarker_groups = compile_biomarker_groups_table()

biomarker_varnames = 
  biomarker_groups |> 
  pull("biomarker")

biomarker_levels = 
  v1_usable |> 
  dplyr::select(all_of(biomarker_varnames)) |> 
  lapply(F = levels)

df = v1_usable

biomarker_events_table =
  construct_biomarker_events_table(
    biomarker_levels,
    biomarker_groups)

nlevs = 
  biomarker_levels |> sapply(length)

```

### Biomarkers of neurodegenerative events

We analyzed `r length(biomarker_levels)` biomarkers (@tbl-biomarker-list) with a total of `r nrow(biomarker_events_table)` non-baseline levels; each level above baseline constitutes an outcome event in the disease progression modelling analysis (@sec-Statistical-analysis).
@tbl-tremors - @tbl-autoimmune summarize the biomarker data, stratified by CGG (`r unique(v1_usable$"FX3*")`) 
and Sex (`r unique(v1_usable$Gender)`).

## Statistical analysis {#sec-Statistical-analysis}

We applied  the Ordinal **Su**btype and **St**age **In**ference ("SuStaIn") algorithm [@young2018uncovering; @young2021ordinal] to our dataset to find event sequences and subtypes for FXTAS patients. This algorithm combines disease progression modelling [@fonteijn2012event] and unsupervised clustering [@lazarsfeld1950logical; @mclachlan2019finite] to model event onset sequences using a cross-sectional sample of a patient population. The algorithm simultaneously clusters individuals into latent subtypes and characterizes the trajectory that best defines each subtype, thus capturing heterogeneity in both disease subtype and disease stage.
Ordinal SuStaIn uses Markov Chain Monte Carlo (MCMC) sampling to estimate the Bayesian posterior probability of each possible event sequence for each subtype given the training dataset, assuming a uniform prior distribution over the set of all possible patterns. 

We fit models stratified by sex, with no latent subtypes (@fig-first-only). We also fit the model stratified by CGG repeats (<100 vs $\ge$ 100), also with no latent subtypes (@fig-pvd-by-cgg).
We then fit the models stratified by both sex and CGG repeats (<100 vs $\ge$ 100), also with no latent subtypes (@sec-stratified-by-cgg-and-sex).

### Incomplete data

```{r}
x1 = v1_usable_cases |> pull(`Ataxia: severity`)
n_missing = sum(x1 |> is.na())
n_obs = sum(x1 |> is.na() |> not()) 
n_total = length(x1)
pct_missing = scales::percent(n_missing/n_total, accuracy = 0.1)
pct_obs = scales::percent(n_obs/n_total, accuracy = 0.1)
probs = x1 |> table() |> proportions() |> as.vector() |> scales::percent(n_obs/n_total, accuracy = 0.01) |> 
  and::and()
```

We handled missing outcome event data by assigning a probability distribution across the possible values of the missing variable that matched the marginal distribution among the cases. 
For example, `r n_missing` of the `length(x1)` cases (`r pct_missing`) had missing values for Ataxia Severity, and `r n_obs` (`r pct_obs`) had recorded values, distributed among severity levels 0-4 ( @tbl-ex-incomplete-data). 
For the `r n_missing` cases with missing values, we assigned probabilities of `r probs` to levels `r and::and(0:4)`, respectively.

```{r}
#| label: tbl-ex-incomplete-data
#| tbl-cap: "Distribution of ataxia severity among cases"
table1(
  na.is.category = FALSE,
  data = v1_usable_cases, 
  x = ~ `Ataxia: severity*`)
```


### Testing for differences between covariate subgroup strata

To test for event sequence differences between males and females and between lower premutation levels (CGG <100) and higher premutation levels (CGG 100-199), we implemented a permutation test, in which we created 1000 permuted datasets in which the strata variables were randomly shuffled. 
We computed the mean log-likelihood of the data for each permuted dataset (averaging across MCMC samples and summing across the strata being compared) and compared the distribution of permuted mean log-likelihoods to the observed log-likelihood calculated from the original (unpermuted) dataset.

### Latent subtypes

We also fit the model on the full dataset (not stratified by sex or CGG) for 2-`r N_S_max` latent subtypes, each with their own ordering (@fig-pvd_optimal). We determined the optimal number of latent subtypes for this dataset using the CVIC criterion (@sec-cvic). More details are provided in @sec-subtypes.

### Longitudinal subsets

We also fit the Ordinal SuStaIn model, stratified by sex, including all longitudinal repeated measurements of the same participants over time (@sec-longitudinal-results). Including later observations might help the model clarify the later stages of the event sequence. However, correlation between repeated observations of the same individuals are not accounted for by the Ordinal SuStaIn model, which assumes that all observations are statistically independent, so including repeated measurements will lead to exaggerated estimates of precision.

### Visualizing modeling results

We visualized the results of Ordinal SuStaIn analysis using "positional variance diagrams" (PVDs), which are heatmaps with biomarker events on the y-axis and sequence positions on the x-axis. 

The PVD's color scale indicates the Bayesian posterior probability that a particular biomarker event (y-axis) appears at a particular position along the progression sequence (x-axis). Red indicates a more probable sequence position, and gray indicates a less probable position.

```{r}
#| label: get-control-dists

control_data = 
  df |> 
  filter(`FX*` == "CGG < 55") |> 
  select(all_of(biomarker_varnames))

patient_data = 
  df |> 
  # na.omit() |>
  filter(`FX*` == "CGG >= 55")

prob_correct = 
  control_data |> 
  compute_prob_correct(
    max_prob = .95,
    biomarker_levels = biomarker_levels)

```

```{r "run OSA from R"}
#| message: false
#| label: model-all-data
#| include: false
#| eval: !expr fit_models

sustain_output = run_and_save_OSA(
  biomarker_levels = biomarker_levels, 
  prob_correct = prob_correct, 
  
  SuStaInLabels = biomarker_varnames,
  N_startpoints = N_startpoints,
  N_S_max = N_S_max, 
  N_iterations_MCMC = N_iterations_MCMC, 
  output_folder = output_folder, 
  dataset_name = dataset_name, 
  use_parallel_startpoints = FALSE,
  seed = 1,
  plot = TRUE,
  patient_data = patient_data,
  N_CV_folds = N_CV_folds)

```

```{r}
#| message: false
#| label: model-males
#| include: false
#| eval: !expr fit_models

sustain_output_males = run_and_save_OSA(
  biomarker_levels = biomarker_levels,
  prob_correct = prob_correct,
  patient_data = patient_data |> filter(Gender == "Male"),
  
  SuStaInLabels = biomarker_varnames,
  N_startpoints = N_startpoints,
  N_S_max = N_S_max_stratified,
  N_iterations_MCMC = N_iterations_MCMC,
  output_folder = output_folder,
  dataset_name = "males",
  use_parallel_startpoints = FALSE,
  seed = 1,
  plot = TRUE,
  rerun = TRUE,
  fig_size = c(20, 10)
)

```


```{r}
#| message: false
#| label: model-females
#| include: false
#| eval: !expr fit_models

sustain_output_females = run_and_save_OSA(
  biomarker_levels = biomarker_levels, prob_correct = prob_correct, 
  patient_data = patient_data |> filter(Gender == "Female"),
  
  SuStaInLabels = biomarker_varnames,
  N_startpoints = N_startpoints,
  N_S_max = N_S_max_stratified, 
  N_iterations_MCMC = N_iterations_MCMC, 
  output_folder = output_folder, 
  dataset_name = "females", 
  use_parallel_startpoints = FALSE,
  seed = 1,
  plot = TRUE,
  rerun = TRUE,
  fig_size = c(20,10))

```


```{r}
#| message: false
#| label: "cgg_over_100"
#| include: false
#| eval: !expr fit_models

sustain_output_cgg100plus = run_and_save_OSA(
  biomarker_levels = biomarker_levels, prob_correct = prob_correct, 
  patient_data = patient_data |> filter(`CGG` >= 100),
  
  SuStaInLabels = biomarker_varnames,
  N_startpoints = N_startpoints,
  N_S_max = N_S_max_stratified, 
  N_iterations_MCMC = N_iterations_MCMC, 
  output_folder = output_folder, 
  dataset_name = "over100", 
  use_parallel_startpoints = FALSE,
  seed = 1,
  plot = FALSE)

```

```{r}
#| message: false
#| label: "cgg_under_100"
#| include: false
#| eval: !expr fit_models
sustain_output_cgg100minus = run_and_save_OSA(
  biomarker_levels = biomarker_levels, 
  prob_correct = prob_correct, 
  patient_data = patient_data |> filter(`CGG` < 100),
  
  SuStaInLabels = biomarker_varnames,
  N_startpoints = N_startpoints,
  N_S_max = N_S_max_stratified, 
  N_iterations_MCMC = N_iterations_MCMC, 
  output_folder = output_folder, 
  dataset_name = "under100", 
  use_parallel_startpoints = FALSE,
  seed = 1,
  plot = FALSE)

```

```{r}
#| label: save-image
fs::path_package("fxtas", "extdata") |> 
  fs::path("ordinal-sustain-results.rda") |> 
  save.image()

# fs::path_package("fxtas", "extdata") |>
#   fs::path("ordinal-sustain-results.rda") |>
#   load()

```

```{r}
#| label: extract-figs

results_v1 = extract_results_from_pickles(
  n_s = 1:N_S_max,
  rda_filename = "data.RData",
  dataset_name = "sample_data",
  output_folder = "output/output.fixed_CV")

fig_females_first = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "females",
  output_folder = "output/output.fixed_CV")

fig_males_first = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "males",
  output_folder = "output/output.fixed_CV")

fig_both_first = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "sample_data",
  output_folder = "output/output.fixed_CV")

fig_under100 = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  dataset_name = "under100",
  output_folder = output_folder)

fig_over100 = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  dataset_name = "over100",
  output_folder = output_folder)

fig_under100_males = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  dataset_name = "under100_Male",
  output_folder = output_folder)

fig_over100_males = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  dataset_name = "over100_Male",
  output_folder = output_folder)


fig_under100_females = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  dataset_name = "under100_Female",
  output_folder = output_folder)

fig_over100_females = extract_figs_from_pickle(size.y = size.y,
  n_s = 1,
  dataset_name = "over100_Female",
  output_folder = output_folder)

```


# Results

## Stratified by sex

```{r}
#| label: p-val-test-sex
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  permuting_variables = "Gender")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('females', 'males'))

pval_sex_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats
)

```

@fig-first-only-1 and @fig-first-only-2 show estimated event sequences stratified by sex. We found statistically significant evidence of a difference in event sequences between males and females (p = `r pval_sex_gp34_v1`). 
It appears that males develop cerebral, cerebellar, and splenium white matter hyperintensity at an earlier stage of the event sequence than females (events 6, 7, and 8 for males versus 9, 1)
ataxia severity earlier in the event sequence than females, while females experience MRI events earlier in the event sequence than males.^[There are probably other patterns worth noting here; clinical folks, please add your observations from the results here.]

## Stratified by CGG repeats

```{r}
#| label: p-val-test-cgg
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  permuting_variables = "FX3")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100', 'over100'))

pval_cgg_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats
)
# autoplot(pval_sex_gp34_v1)
```

We did not find statistically significant evidence of a difference in event sequences between CGG < 100 and CGG > 100,  (p = `r pval_cgg_gp34_v1`).
@fig-pvd-by-cgg shows the estimated sequence models stratified by CGG repeat level.


## Stratified by CGG and Sex {#sec-stratified-by-cgg-and-sex}

```{r}
#| label: p-val-test-cgg-male
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "Male",
  permuting_variables = "FX3")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100_Male', 'over100_Male'))

pval_male_cgg_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats
)

# autoplot(pval_male_cgg_gp34_v1)

```

We found significant evidence of a difference between CGG < 100 and CGG ≥ 100 among males (p = `r pval_male_cgg_gp34_v1`). @fig-pvd-by-cgg-male shows the corresponding stratified models.

```{r}
#| label: p-val-test-cgg-female
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "Female",
  permuting_variables = "FX3")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100_Female', 'over100_Female'))

pval_female_cgg_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats
)

```

We did not find statistically significant evidence of a difference between CGG < 100 and CGG ≥ 100 among females (p = `r pval_female_cgg_gp34_v1`). @fig-pvd-by-cgg-female shows the estimated stratified models.

```{r}
#| label: p-val-test-sex-under100
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "CGG 55-99",
  permuting_variables = "Gender")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100_Female', 'under100_Male'))

pval_under100_sex_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats
)

# autoplot(pval_under100_sex_gp34_v1)

```

We did not find statistically significant evidence of a difference between males and females among those with CGG < 100 (p = `r pval_under100_sex_gp34_v1`). @fig-pvd-by-gender-cgg_under100 shows the estimated stratified models.

```{r}
#| label: p-val-test-sex-over100
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "CGG 100-199",
  permuting_variables = "Gender")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('over100_Female', 'over100_Male'))

pval_over100_sex_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats
)

# autoplot(pval_over100_sex_gp34_v1)

```

We did not find statistically significant evidence of a difference between males and females among those with CGG < 100 (p `r pval_over100_sex_gp34_v1`). @fig-pvd-by-gender-cgg_over100 shows the estimated stratified models.

# Discussion

## Missing data

We had substantial amounts of missing data. As described in the methods, we used a uniform distribution to impute the underlying values of these missing data. The missing data contributes to the large amount of uncertainty in the results.

[to be added]

# References {.unnumbered}

::: {#refs}
:::


# Tables {.unnumbered}

```{r}
#| tbl-cap: "Biomarkers used in analysis"
#| label: "tbl-biomarker-list"

probs_above_baseline_by_gender = 
  v1_usable |> 
  summarize(
    .by = "Gender",
    across(
      all_of(biomarker_varnames),
      ~ mean(.x != levels(.x)[1], na.rm = TRUE)
           )
  ) |> 
  pivot_longer(
    cols = -Gender, 
    names_to = "biomarker", 
    values_to = "Pr(above_baseline)") |>
  mutate(
    `Pr(above_baseline)` = 
      `Pr(above_baseline)` |>  
      scales::percent(accuracy = 0.01)) |> 
  pivot_wider(
    id_cols = "biomarker",
    names_from = Gender,
    values_from = `Pr(above_baseline)`)

table_out =
  biomarker_events_table |> 
  select(
    category = biomarker_group, 
    biomarker, 
    levels) |> 
  slice_head(by = biomarker) |> 
  filter(category != "Stage") |> 
  left_join(
    probs_above_baseline_by_gender,
    by = "biomarker",
    relationship = "one-to-one"
  ) |> 
  mutate(
    biomarker = 
      biomarker |> 
      sub(
        pattern = "*", 
        replacement = "", 
        fixed = TRUE))

table_out |> pander()

```

::: landscape

```{r}
#| label: "tbl-demographics"
#| tbl-cap: "Demographics by CGG repeats level"

vars = c(
  # "Study",
  "Age at visit",
  # "# visits",
  # column_var,
  "Primary Race/Ethnicity",
  "Primary Ethnicity",
  "Primary Race",
  "FXTAS Stage (0-5)*"
  # "ApoE"
  
)

# table1(
#   na.is.category = FALSE, overall = FALSE,
#   NA.label = "Missing",
#   stratified_formula(vars, strata = "FX*"),
#   # render.continuous = c(.="N", .="Mean (SD)", .="Median [Min,  Max]"),
#   data = v1_usable)


# create table using gtsummary with p-value for sex difference
tbl_stat <- tbl_strata(
  data = v1_usable |> clean_ethnicity() |> create_race_ethnicity() |> 
    dplyr::select(all_of(vars), Gender, `FX*`),
  strata = c("Gender"),
  .tbl_fun = ~ .x %>%
    tbl_summary(
      by = `FX*`,
      type = all_continuous() ~ "continuous2",
      statistic = list(
        all_continuous() ~ c("{mean} ({sd})", "{median} [{min}, {max}]")
      ),
      digits = list(`Age at visit` ~ c(1, 2)),
      missing_text = "Missing"
    )
)

tbl_pval <- v1_usable |> clean_ethnicity() |> create_race_ethnicity() |>
  dplyr::select(all_of(vars), Gender) |>
  tbl_summary(
    by = Gender,
    missing_text = "Missing"
  ) |>
  add_p(
    pvalue_fun = function(x) style_number(x, digits = 3)
  ) |>
  modify_column_hide(columns = c(stat_1, stat_2))

tbl_final <- tbl_merge(
  list(tbl_stat, tbl_pval), tab_spanner = FALSE
)

tbl_final
```

{{< pagebreak >}}

# Figures {.unnumbered}

```{r}
#| label: "fig-first-only"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequence models, stratified by sex, first visits only"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "Males"
#|   - "Females"


# fig_both_first |> print()
fig_males_first |> print()
fig_females_first |> print()

```

{{< pagebreak >}}

```{r}
#| label: "fig-pvd-by-cgg"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences, stratified by CGG repeats (<100 vs 100+); first visits only"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "CGG < 100"
#|   - "CGG 100+"

# fig_both_first |> print() 
fig_under100 |> print()
fig_over100 |> print()

```

{{< pagebreak >}}

```{r}
#| label: "fig-pvd-by-cgg-male"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences, stratified by CGG repeats (<100 vs 100+); first visits only, males only"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "Males, CGG < 100"
#|   - "Males, CGG ≥ 100"

# fig_males_first |> print() 
fig_under100_males |> print()
fig_over100_males |> print()

```

{{< pagebreak >}}

```{r}
#| label: "fig-pvd-by-cgg-female"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences, stratified by CGG repeats (<100 vs 100+); first visits only, females only"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "Females, CGG < 100"
#|   - "Females, CGG ≥ 100"

# fig_females_first |> print() 
fig_under100_females |> print()
fig_over100_females |> print()

```

{{< pagebreak >}}

```{r}
#| label: "fig-pvd-by-gender-cgg_under100"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences, stratified by sex, for CGG repeats <100; first visits only"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "Males, CGG < 100"
#|   - "Females, CGG < 100"

fig_under100_males |> print()
fig_under100_females |> print()

```

{{< pagebreak >}}

```{r}
#| label: "fig-pvd-by-gender-cgg_over100"
#| column: page
#| fig-height: !expr pvd_height
#| fig-cap: "Estimated event sequences, stratified by sex, for CGG repeats ≥100; first visits only"
#| fig-cap-location: top
#| layout-ncol: 2
#| fig-subcap:
#|   - "Males, CGG ≥ 100"
#|   - "Females, CGG ≥ 100"

fig_over100_males |> print()
fig_over100_females |> print()

```

:::

# Supplementary material {.appendix}

{{< pagebreak >}}

## Distribution of CGG repeats

```{r}
#| label: fig-hist-CGG-gp34-v1
#| fig-cap: "Distribution of CGG repeats by sex, among study participants included in analysis"
#| warning: false

v1_usable |> graph_CGG_repeats(rows = vars(Gender))
```

{{< pagebreak >}}

{{< include general-data-notes.qmd >}}

## Hypothesis testing {.appendix}

@fig-permute-llik-sex compares the observed log-likelihood for the sex-stratified model to the distribution of mean log-likelihoods when sex is permuted.

{{< pagebreak >}}

```{r}
#| fig-cap: "Histogram of mean log-likelihoods for GP34 visit 1 data, permuted by sex"
#| label: fig-permute-llik-sex

pval_sex_gp34_v1 |> autoplot() |> print()
```

{{< pagebreak >}}

## Results with repeated measures {#sec-longitudinal-results .appendix}

{{< include longitudinal-results.qmd >}}

## Detecting latent subtypes {#sec-subtypes .appendix}

In order to use the Ordinal SuStaIn modeling algorithm, we must specify how many latent subtypes to include in the model. There are several metrics for determining the optimal number of subtypes for a given data set.

### Likelihood {.appendix}

The simplest option is to compare the likelihood of the data for the fitted model while varying the number of subtypes used to fit the model. @fig-mcmc-loglik shows the distribution of log-likelihoods from the MCMC samples for the full dataset (not stratified by sex or CGG repeats). Adding up to 6 clusters substantially improves the log-likelihood.

```{r}
#| fig-height: 4
#| fig-width: 6
#| label: fig-mcmc-loglik
#| fig-cap: log-likelihoods of MCMC samples, by number of subtypes
#| eval: !expr fit_models

lliks = results_v1 |> 
  sapply(F = function(x) x$samples_likelihood)

lliks |>
  graph_likelihoods_v2(alpha = 0.5) |>
  suppressWarnings()

```

{{< pagebreak >}}

### Cross-Validation Information Criterion {#sec-cvic .appendix}

Since models with fewer subtypes are limited special cases of models with more subtypes (i.e., "nested models"), the likelihood of the training data is guaranteed to improve as we increase the number of subtypes. However, using too many subtypes means there is less data available for each subtype, possibly leading to a model that does not generalize well to new data (i.e., a model that is "overfit" to the training data).

To avoid overfitting, @young2018uncovering recommends choosing the optimal number using the Cross-Validation Information Criterion (CVIC):

#### Mathematical details of cross-validation criteria {#sec-cvic-math .appendix}

Let:

* $n$ be the number of observations in the dataset
* $Y_i$ be the observed biomarker values for the $i^{th}$ participant
* $\mathcal C$ be the number of latent subtypes assumed, with corresponding index $c$.
* $C_i$ be the latent subtype membership of observation $i$
* $\mathcal S$ be the set of possible event sequences, indexed by $\mathcal s$
* $S_c \in \mathcal S$ be the sequence for subtype $c$
* $K$ be the number of cross-validation folds, with index $k$; typically, $K = 10$.
* $n_k$ be the number of held-out observations in cross-validation fold $K$; $n_k \approx n/K$
* $\hat{P}_{(-k)}(C = c)$ be the prior probability that an observation belongs to subtype $c$, estimated using all observations except those in fold $k$.

Then:

$$
CVIC_{\mathcal C} = 
-2 * 
\sum_{k=1}^K 
\sum_{i=1}^{n_k} 
\log
\left\{
\sum_{c=1}^\mathcal C 
\sum_{s\in \mathcal S}
\hat P(Y_i|C_i = c, S_c=s)
\hat{P}_{(-k)}(C_i = c, S_c = s)
\right\}
$$

#### CVIC for the full dataset {.appendix}

We performed `r N_CV_folds`-fold cross-validation on the unstratified data, and calculated the CVIC for 1-`r N_S_max` latent subtypes.

```{r}
#| fig-cap: "Cross-validation information criterion"
#| label: fig-cvic
#| eval: !expr fit_models

temp = sustain_output |> attr("CV")
temp$CVIC |> plot_CVIC()

```

It appears that the model reaches optimum CVIC at 6 subtypes.

## Out-of-fold log-likelihood distribution {.appendix}

We also evaluated the consistency of our cross-validation procedure by looking at the distribution of test-set log-likelihood across folds (@fig-boxplot-loglik-cv); following @young2018uncovering, we calculated out-of-sample log-likelihood as:

$$
\ell_{\mathcal C, k} = 
\sum_{i = 1}^{n_k}
\sum_{c=1}^\mathcal C 
\sum_{s\in \mathcal S}
\log
\left\{
\hat P(Y_i|C_i = c, S_c=s)
\right\}
\hat{P}_{(-k)}(C_i = c, S_c = s)
$$
{{< pagebreak >}}

```{r}
#| fig-cap: "Test set log-likelihood across folds"
#| label: fig-boxplot-loglik-cv
#| eval: !expr fit_models
temp$loglike_matrix |> 
  plot_cv_loglik()

```

{{< pagebreak >}}

{{< include results-latent-subgroups.qmd >}}

::: landscape
## Additional data tables {.appendix}

{{< include exploration.qmd >}}
{{< include data-tables-appendix.qmd >}}
:::
