---
title: "Development and progression of 
Fragile X-associated tremor/ataxia syndrome 
revealed by Subtype and Stage Inference analysis"
format:
  html:
    echo: false
  docx: default
reference-doc: ../Brain_template_2022.docx
---

```{=html}
<style>
.quarto-figure-center > figure {
text-align: center;
}
</style>
```

```{r}
#| include: false
#| label: setup
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  include = TRUE
)
pander::panderOptions("table.split.table", Inf)
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| label: libraries
#| message: false
library(fxtas)
reticulate::use_condaenv("fxtas39", required = TRUE)
library(tidyverse)
library(reticulate)
library(pander)
library(table1)
library(magrittr)
library(knitr)
```


```{r}
#| label: "set run parameters"

max_prob_correct = .95

fit_models <- TRUE
run_cv <- TRUE

rerun  <- TRUE
# rerun <- FALSE

use_rds <- TRUE
use_rds <- FALSE
plot_python <- TRUE
fig_size <- c(20, 10)

pvd_height = 10
pvd_width = 8
pvd_line_height = 10
pvd_line_width = 8
size.y = 11
N_startpoints = 10L
# use_parallel_startpoints = TRUE
use_parallel_startpoints = FALSE # couldn't get parallel to work, 2024/06/15
N_S_max = 8L
N_S_max_stratified = 2L
N_CV_folds = 10L

n_permutations = 1000

outcomes_name = "symptom"
events_name = "symptomatic events"
outcomes_name_sentence_start = stringr::str_to_sentence(outcomes_name)

compact_fig_caption <- paste(
  "**Positional variance diagrams of",
  "estimated event sequences models and uncertainty.**",
  "The different colors (red, blue, purple, green, magenta)",
  "indicate the ordinal levels of symptom progression.",
  "Color intensity represents the likelihood of sequence position.",
  "The brighter the color, the more likely that the corresponding",
  outcomes_name,
  "event occurs in that position in the sequence."
)

compact_fig_cap2 = glue::glue(
  "Red lines indicate {outcomes_name}s that moved to later positions ",
  "between the left-hand subgroup and the right-hand subgroup. ",
  "Blue lines indicate ",
  "{outcomes_name}s that moved to earlier positions. ",
  "Line opacity levels indicate the number of positions changed ",
  "(more opaque = more change)."
)

# column_var = "Recruited in study phase"
# column_var = "FX3*"
column_var = c("Gender", "FX3*")

N_iterations_MCMC = 1e5L
dataset_name = 'sample_data'
root_dir = here::here()
setwd(root_dir)
output_folder =
  "output/output.fixed_CV/" |>
  fs::dir_create()


```

```{r}
#| label: subset-data

full_data = trax_gp34_v1
n_missing_CGG = full_data$CGG |> is.na() |> sum()
n_above_200 = sum(full_data$CGG >= 200, na.rm = TRUE)
v1_usable = 
  full_data |> 
  filter(CGG < 200) |>
  mutate(`FX3*` = `FX3*` |> forcats::fct_drop() |> 
           labelled::set_label_attribute("CGG Repeat Level"),
         Parkinsons = Parkinsons |> 
           labelled::set_label_attribute("Parkinson's disease"),
         `FXTAS Stage (0-5)*` = `FXTAS Stage (0-5)*` |> 
           labelled::set_label_attribute("FXTAS Stage"))

v1_usable_males = v1_usable |> filter(Gender == "Male")
v1_usable_females = v1_usable |> filter(Gender == "Female")
v1_usable_cases = v1_usable |> filter(CGG >= 55)
v1_usable_male_cases = 
  v1_usable |> 
  filter(Gender == "Male", 
         CGG >= 55)

v1_usable_female_cases = 
  v1_usable |> 
  filter(Gender == "Female", 
         CGG >= 55)

v1_usable_male_controls = 
  v1_usable |> 
  filter(Gender == "Male", 
         CGG < 55)

v1_usable_female_controls = 
  v1_usable |> 
  filter(Gender == "Female", 
         CGG < 55)
# note: there are 231 records in `visit1` with CGG >= 55, but 4 have CGG >= 200
# previously `nrow(v1_usable_cases)` was 221, 
# which was based on incorrectly filtering on a version of CGG 
# that hadn't been backfilled.

controls_v1 =  v1_usable |> dplyr::filter(CGG < 55)
year_range = v1_usable$`Visit Date` |> year() |> range()
min_year = min(year_range)
max_year = max(year_range)

control_data = controls_v1
patient_data = v1_usable_cases

```

```{r}
#| label: "biomarker-events"
biomarker_groups = compile_biomarker_groups_table(
  dataset = v1_usable
)

biomarker_varnames = 
  biomarker_groups |> 
  pull("biomarker")

biomarker_levels = 
  v1_usable |> 
  dplyr::select(all_of(biomarker_varnames)) |> 
  lapply(F = levels)

biomarker_events_table =
  construct_biomarker_events_table(
    biomarker_levels,
    biomarker_groups = biomarker_groups)

nlevs = 
  biomarker_levels |> sapply(length)

```

[Target journal: 
[Brain (Original Articles)](https://academic.oup.com/brain/pages/general_instructions).
Word limit for main text: 6000.
Figure + table limit: 8.]{.mark}

{{< pagebreak >}}

# Abstract {.unnumbered}

[Abstract word limit: 400]{.mark}

{{< include _sec_abstract.qmd >}}

# Introduction

{{< include _sec_intro.qmd >}}

# Materials and methods

## Data

### Study cohorts

{{< include _sec_study_cohorts.qmd >}}

### `r outcomes_name_sentence_start`s of neurodegenerative events

{{< include _sec_biomarkers.qmd >}}

## Statistical analysis {#sec-Statistical-analysis}

### Ordinal SuStaIn model {#sec-scored-events-model}

{{< include _sec_ordinal_sustain_model.qmd >}}

### Imputation of missing data {#sec-incomplete-data}

{{< include _sec_incomplete_data.qmd >}}

### Statistical hypothesis tests

{{< include _sec_hypothesis_tests.qmd >}}

### Latent subtype clustering {#sec-methods-latent-subtypes}

We also fitted the model on the full dataset (not stratified by sex or CGG) 
for 2 - `r N_S_max` latent subtypes, 
each with their own ordering. 
We determined the optimal number of latent subtypes for this dataset using 
the Cross-Validation Information Criterion (CVIC)
and the Out-Of-Fold Log-Likelihood (OOFLL) criterion [@young2018uncovering];
both criteria quantify how well models containing a given number of subtypes
extend to new data not used in training.
We performed `r N_CV_folds`-fold cross-validation on the unstratified data
and calculated the CVIC and OOFLL for 1-`r N_S_max` latent subtypes. 

<!-- {{< include _methods_longitudinal.qmd >}} -->

### Visualizing modeling results

We visualized the results of Ordinal SuStaIn analysis 
using "positional variance diagrams" (PVDs).
PVDs are heatmaps with `r events_name` on the y-axis 
and sequence positions of events on the x-axis.
Each event is estimated to occur at some point in continuing time 
*relative to* other events in sequence.
It should be noted that the exact onset time of each event
cannot be determined by the method we used for this analysis.
The PVD's color scale indicates the Bayesian posterior probability that 
a particular event (y-axis) 
appears at a particular position along the progression sequence (x-axis). 
The different colors (red, blue, purple, green, magenta)
indicate the ordinal levels of symptom progression.
Color intensity presents the likelihood of sequence position. 
That is, a brighter color indicates a more probable sequence position, 
and a paler color indicates a less probable position. 

## Software

The Ordinal Sustain analysis was performed using 
the Python programming language, version 3.9 [@python3] 
with the `pySuStaIn` package [@aksman2021pysustain].
Data pre-processing and results post-processing were performed in 
R 4.4.0 [@rcore] using the `tidyverse` packages [@tidyverse].
The `reticulate` package [@reticulate] was used to create 
an application programming interface between Python and R.
The code used to perform this analysis is available at 
<https://github.com/d-morrison/fxtas>.

# Results

```{r}
#| label: get-control-dists

prob_correct = 
  control_data |> 
  compute_prob_correct(
    max_prob = max_prob_correct,
    biomarker_levels = biomarker_levels)

```

```{r}
#| message: false
#| label: model-all-subgroups-first-visits
#| include: false
#| eval: !expr fit_models

sustain_output = run_and_save_OSA(
  biomarker_levels = biomarker_levels, 
  prob_correct = prob_correct, 
  SuStaInLabels = biomarker_varnames,
  N_startpoints = N_startpoints,
  N_S_max = N_S_max, 
  N_iterations_MCMC = N_iterations_MCMC, 
  output_folder = output_folder, 
  dataset_name = dataset_name, 
  use_parallel_startpoints = use_parallel_startpoints,
  seed = 1,
  plot = plot_python,
  rerun = rerun,
  patient_data = patient_data,
  N_CV_folds = N_CV_folds)

```

{{< include _sec_run_stratified_models.qmd >}}

```{r}
#| label: extract-figs

results_v1 = extract_results_from_pickles(
  use_rds = use_rds,
  n_s = 1:N_S_max,
  rda_filename = "data.RData",
  dataset_name = "sample_data",
  output_folder = "output/output.fixed_CV")

fig_females_first = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "females",
  output_folder = "output/output.fixed_CV")

fig_males_first = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "males",
  output_folder = "output/output.fixed_CV")

fig_both_first = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  rda_filename = "data.RData",
  dataset_name = "sample_data",
  output_folder = "output/output.fixed_CV")

fig_under100 = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  dataset_name = "under100",
  output_folder = output_folder)

fig_over100 = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  dataset_name = "over100",
  output_folder = output_folder)

fig_under100_males = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  dataset_name = "under100_Male",
  output_folder = output_folder)

fig_over100_males = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  dataset_name = "over100_Male",
  output_folder = output_folder)


fig_under100_females = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  dataset_name = "under100_Female",
  output_folder = output_folder)

fig_over100_females = extract_figs_from_pickle(
  use_rds = use_rds,
  size.y = size.y,
  n_s = 1,
  dataset_name = "over100_Female",
  output_folder = output_folder)

```

```{r}
#| label: basic-summary-stats
stages = v1_usable_cases |> select(`FXTAS Stage (0-5)*`) |> table()
biomarkers_table = 
  v1_usable |> 
  make_biomarkers_table(
    biomarker_varnames = biomarker_varnames,
    biomarker_events_table = biomarker_events_table)
```

@tbl-demographics describes summary statistics of patient characteristics 
included in the analysis. 
Our data included: 
`r nrow(v1_usable_cases)` Fragile X premutation carriers and 
`r nrow(controls_v1)` controls;
`r nrow(v1_usable_males)` males and `r nrow(v1_usable_females)` females;
and `r and::and(stages)` carriers at 
FXTAS stages `r names(stages) |> and::and()`, respectively. 
We found no differences between males and females in 
age at baseline visit, 
ethnicity/race, 
FXTAS stage, or
CGG repeat length.
@tbl-biomarker-list lists the `r outcomes_name`s and their ordinal levels
and reports proportions of clinically elevated levels at baseline visit, 
stratified by sex.
We found significant sex differences in 
proportions of clinically elevated (non-reference) `r outcomes_name` levels 
at baseline visit for `r report_sex_differences(biomarkers_table)`.

## Sex differences in sequential orders {#sec-methods-strat-sex}

```{r}
#| label: p-val-test-sex
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  permuting_variables = "Gender")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('females', 'males'))

pval_sex_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

```

@fig-first-only-pvd shows 
the estimated sequential orders
of FXTAS `r outcomes_name`s 
in males and females, 
and @fig-first-only-diffs shows differences in those orders.
We found statistically significant evidence of a 
difference in sequences of `r events_name`
between males and females (p = `r pval_sex_gp34_v1`). 
It appears that 
female premutation carriers developed 
SCID mood disorder symptoms, 
both sub-threshold-level and threshold-level,
prior to FXTAS Stage 1 diagnosis,
whereas male carriers experienced these symptoms 
much later (between FXTAS Stages 3 and 4) (@fig-first-only-diffs).
Ataxia occurred prior to Stage 1 for females 
and after Stage 2 for males.
However, ataxia severity level 1 occurred 
earlier for males (between Stages 2-3) 
than for females (between Stages 3-4).
Advanced degrees of ataxia severity also occurred 
earlier in the event sequence for males than females. 
MRI biomarkers of decline and parkinsonian features 
also appear to begin at earlier FXTAS stages for males than for females.


```{r}
#| label: p-val-test-sex-under100
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "CGG 55-99",
  permuting_variables = "Gender")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100_Female', 'under100_Male'))

pval_under100_sex_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

```

```{r}
#| label: p-val-test-sex-over100
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "CGG 100-199",
  permuting_variables = "Gender")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('over100_Female', 'over100_Male'))

pval_over100_sex_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

```

When we conducted subgroup analyses by CGG repeat size,
we found statistically significant evidence of 
a difference between males and females
among those with CGG < 100 (p = `r pval_under100_sex_gp34_v1`)
(@suppfig-pvd-by-gender-cgg_under100).
Several psychiatric disorders (as assessed by SCID) 
occurred prior to FXTAS Stage 1
among females with CGG < 100,
but at later FXTAS stages for males with CGG < 100.
MRI biomarkers appeared to occur at later stages in females compared to males.
We did not find statistically significant evidence of 
a difference between males and females
among those with CGG ≥ 100 (p = `r pval_over100_sex_gp34_v1`) 
(@suppfig-pvd-by-gender-cgg_over100).


```{r}
#| label: p-val-test-cgg-male
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "Male",
  permuting_variables = "FX3")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100_Male', 'over100_Male'))

pval_male_cgg_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

```


```{r}
#| label: p-val-test-cgg-female
output_folder = "output/output.fixed_CV/"

permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "Female",
  permuting_variables = "FX3")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100_Female', 'over100_Female'))

pval_female_cgg_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

n_over100_Female =
  v1_usable |>
  filter(
   Gender == "Female",
   CGG >= 100
  ) |>
  nrow()

```

We found significant evidence of an overall difference between CGG <100
and CGG ≥100 among males (p = `r pval_male_cgg_gp34_v1`) 
(@suppfig-pvd-by-cgg-male). 
SWM between errors occurred prior to Stage 1 in males with CGG repeats <100, 
whereas it occurred between stage 3 and 4 
in males with CGG repeats ≥ 100 
([Supplementary Figure @fig-pvd-by-cgg-male-2]). 
Postural, 
resting, 
and intermittent tremor 
occurred in earlier FXTAS stages in those with CGG repeats <100 
compared to those with CGG repeats ≥100.
Several psychiatric disorders (as measured by SCID) 
occurred later in the event sequence in participants with CGG repeats <100 
compared to those with CGG repeats ≥100.  
However, we did not find a significant difference between CGG <100 and CGG ≥100 
among females (p = `r pval_female_cgg_gp34_v1`) 
(@suppfig-pvd-by-cgg-female). 
It should be noted that this comparison lacked statistical power 
given the limited sample size of only `r n_over100_Female` females with CGG ≥100.

## Subtype clustering

```{r}
#| label: set-optimal-subtypes
temp = sustain_output |> attr("CV")
n_s_min_CVIC = which.min(temp$CVIC)
n_s_selected = 4

results_cv_max = extract_results_from_pickle(
  n_s = n_s_selected,
  dataset_name = dataset_name,
  output_folder = output_folder)
```

We conducted latent subtype classification analysis to cluster participants 
that are relatively homogeneous within the same cluster 
and heterogeneous from other clusters 
based on similarities and difference in sequential patterns. 
Based on the CVIC criterion (Supplementary Materials, @sec-subtypes), 
`r n_s_min_CVIC` subtypes were suggested to be the optimal number for the full, 
unstratified dataset 
([Supplementary Figure @fig-cvic-1]). 
The out-of-fold log-likelihood ("OOFLL") (Supplementary Materials, @sec-subtypes) 
showed substantial fold-to-fold variation 
([Supplementary Figure @fig-cvic-2]).
Between `r broman::spell_out(n_s_selected)` and `r broman::spell_out(N_S_max)` 
latent subgroups, 
the distribution of OOFLL appears to be
approximately unchanging and the CVIC appears approximately flat, 
and thus for easier clinical interpretation, 
we chose to classify participants into `r n_s_selected` subtypes.
The subtypes are numbered 1-4 in order of estimated relative frequency
(that is, in order of how many participants were clustered into each subtype).

```{r}
#| label: compute-n-in-subtype-0
n_type_0 = results_cv_max$subtype_and_stage_table |> 
  filter(ml_subtype == "Type 0") |> 
  nrow()

```

@fig-pvd_chosen_subtypes shows the estimated sequential event orders
for each of the subtype clusters.
@suppfig-pvd_chosen_subtypes_linegraphs shows 
differences in event sequences between subtypes. 
@tbl-sg_demos shows 
the demographics of the patients clustered in each subtype. 
`r n_type_0 |> xfun::numbers_to_words() |> Hmisc::capitalize()` 
patients had experienced too few events 
to be accurately classified into a subtype 
and were excluded from @tbl-sg_demos.
Subtype clustering analysis showed that 
subtypes were mainly classified by sex and CGG repeat size, 
supporting the results above. 
Subtype 1 appeared more biased towards males (65.1%) 
and lower CGG repeat size (83.2 ± 15.8) 
compared to the overall study population 
(64.5% of males; CGG repeat size = 87.3  ± 18.9) and other subtypes; 
Subtype 2 appeared biased towards females (53.1% of males) 
and higher CGG repeat size (90.2 ± 22.3); 
Subtype 3 appeared over-represented by males (82% of males) 
and high CGG repeat size (92.4 ± 17.4); 
and Subtype 4 was biased towards females (60% of males) 
and lower CGG repeat (82.7 ± 17.3). 

```{r}
n_events = nrow(biomarker_events_table)
```

Using the FXTAS Stage transitions as reference points,
we examined the differences in the appearing symptoms
relative to these FXTAS Stage transitions.
Roughly speaking:
Subtype 1 appeared to experience 
early mood disorders (before FXTAS Stage 1),
ataxia between Stages 2 and 3,
head tremors between Stages 3 and 4,
parkinsonian features between Stages 3 and 4,
and white matter disease between Stages 4 and 5.
Compared to Subtype 1, Subtype 2 appeared to experience 
white matter disease phenotypes earlier (between Stages 2 and 4),
memory impairment (assessed by SWM) earlier (Stage 3-4 versus 4-5),
and psychiatric disorders later (between Stages 3 and 5).
Subtype 2 might be described as "memory-first".
Subtype 3 appeared to experience
white matter disease phenotypes early (between Stages 1 and 3),
and mood disorders late (Stages 4-5).
Subtype 4 had the fewest participants and was difficult to interpret.
It may represent a mixture of smaller latent subtypes and/or
include outlier individuals with unusual event sequences.

# Discussion

{{< include _sec_discussion.qmd >}}

# Data availability {.unnumbered}

The de-identified data used in this analysis may be provided 
upon request 
from the principal investigators of the two cohorts.

# Funding {.unnumbered}

This study was supported by the 
National Institute of Child Health and Human Development 
(NICHD, HD036071) 
and
UC Davis MIND Institute’s 
Intellectual and Developmental Disabilities Research Center (IDDRC) 
(P50HD103256)
and the National Institute of Neurological Disorders and Stroke 
(NINDS, NS110100).

# Competing interests {.unnumbered}

The authors report no competing interests.

# Supplementary material {.unnumbered}

Supplementary material is available at Brain online.

# References {.unnumbered}

::: {#refs}
:::

::: landscape

# Tables {.unnumbered}

```{r}
#| label: "tbl-demographics"
#| tbl-cap: |
#|    **Descriptive statistics of patient characteristics by CGG repeat level.**
#|    `*`p-values represent tests for sex differences in distributions of 
#|    characteristics, all CGG repeat levels combined.

v1_usable |> make_demographics_table()

```

{{< pagebreak >}}

```{r}
#| tbl-cap: |
#|    **Symptoms included in analysis.** 
#|    The first level listed in each row of column "Defined Ordered Levels" 
#|    is the "reference level" for the corresponding symptom,
#|    and subsequent levels are considered "clinically elevated levels".
#|    Columns "Female" and "Male" list proportions of clinically elevated levels 
#|    at baseline visit, stratified by sex.
#| label: tbl-biomarker-list

flex_biomarkers_table(biomarkers_table)
```

:::

```{r}
#| label: build-tbl-sg-demos
set.seed(2)
tbl_sg_demos = table_subtype_by_demographics(
  patient_data,
  subtype_and_stage_table =
    results_cv_max$subtype_and_stage_table,
  digits = list(
    CGG = 1,
    `FX3*` = c(0, 1),
    Gender = c(0, 1),
    `Primary Race/Ethnicity` = c(0, 1)
  )
)
```

```{r}
#| tbl-cap: !expr 
#|     glue::glue(
#|        "**Demographics of {n_s_selected} latent subtype clusters 
#|        identified by Ordinal SuStaIn.** 
#|        {n_type_0} patients had experienced too few events 
#|        to be accurately classified into a subtype 
#|        and were excluded from this table.")
#| label: tbl-sg_demos
#| results: asis


tbl_sg_demos |>
  gtsummary::as_flex_table() |>
  flextable::width(width = 1.75) |>
  flextable::width(j = 2:6, width = 0.75) |>
  flextable::width(j = 7, width = 1)

```

{{< pagebreak >}}

# Figures {.unnumbered}

```{r}
#| label: fig-first-only-pvd
#| column: page
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| fig-cap: !expr 
#|    glue::glue(
#|    "**Event sequence models stratified by sex.** 
#|    {compact_fig_caption}")
#| fig-cap-location: top


plot_compact_pvd(
  figs = list(
    fig_males_first,
    fig_females_first
  ),
  facet_label = c("Males <br>", "Females <br>"),
  y_text_size = 11
)

```

{{< pagebreak >}}

```{r}
#| label: fig-first-only-diffs
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| fig-cap: !expr 
#|    glue::glue(
#|    "**Positional differences in estimated event sequence between sexes.** 
#|    {compact_fig_cap2}")
#| fig-cap-location: top

list(
    "Males" = fig_males_first,
    "Females" = fig_females_first
  ) |> 
pvd_lineplot(y_text_size = 11)

```

{{< pagebreak >}}

```{r}
#| label: fig-pvd_chosen_subtypes
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| fig-cap: !expr 
#|     glue::glue(
#|     "**Event sequences for {n_s_selected} latent subtypes.**
#|     {compact_fig_caption}")
#| fig-cap-location: top

figs = extract_figs_from_pickle(
  size.y = size.y,
  n_s = n_s_selected,
  dataset_name = dataset_name,
  output_folder = output_folder)

library(ggplot2)

plot_compact_pvd_est2(
  figs = figs,
  tile_height = 1,
  y_text_size = 8
)


```

{{< pagebreak >}}

# Supplementary Material for “Development and progression of Fragile X-associated tremor/ataxia syndrome revealed by Subtype and Stage Inference analysis" {.appendix}

{{< include _sec_supplement.qmd >}}

