---
title: "potential variables list"
format: 
  html:
    embed-resources: true
editor: visual
---


```{=html}
<style>
.quarto-figure-center > figure {
  text-align: center;
}
</style>
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  # fig.width = 7, 
  # fig.height = 10,
  include = TRUE
)
```

```{r setup}
#| message: false
devtools::load_all()
library(tidyverse)
library(reticulate)
library(pander)
```

```{r}

biomarker_group_list = 
  all_potential_biomarkers(gp34)

biomarker_groups = 
  biomarker_group_list |> 
  compile_biomarker_groups_table()

```

@tbl-biomarker-list lists the biomarkers that are potentially usable for analysis by the Ordinal SuStaIn method.

Notes: 

- "any autoimmune disorder" includes:
  - "Lupus",
  - "Rheumatoid arthritis",
  - "Multiple Sclerosis: Workup",
  - "ANA positive",
  - "Sjogrens Syndrome",
  - "Raynauds Syndrome",
  - "Pulmonary Fibrosis"
  
- "Any tremor (excluding head)" includes:
  - "Intention tremor",
  - "Resting tremor",
  - "Postural tremor",
  - "Intermittent tremor"
  
- "Any cancer" includes:
  - "Thyroid Cancer",
  - "Skin Cancer",
  - "Melanoma",
  - "Prostate Cancer",
  - "Other Cancer"

- an asterisk ("*") after the biomarker name indicates a numerical variable that has been categorized.

```{r}
#| tbl-cap: "Potential biomarkers for Ordinal SuStaIn analysis"
#| label: "tbl-biomarker-list"
table_out =
  biomarker_groups |> 
  dplyr::select(category = biomarker_group, biomarker) |> 
  slice_head(by = biomarker)

table_out |> 
  pander()

writexl::write_xlsx(table_out, 
                    path = fs::path(here::here(), "output", "potential_variables.xlsx"))

```
