
## Hypothesis testing {.appendix #sec-permutation-test}

```{r}
#| label: hypothesis-tests
obs_stat = pval_sex_gp34_v1  |> attr("observed_test_stat")
perm_stats = pval_sex_gp34_v1  |> attr("permuted_test_stats")
n_perm_over_obs = sum(perm_stats >= obs_stat)
pct_perm_over_obs = mean(perm_stats >= obs_stat)
```

@fig-permute-llik-sex illustrates the permutation test approach that we implemented to compare event orderings between subgroups.
This figure compares the mean log-likelihood for the sex-stratified model
to the distribution of mean log-likelihoods when sex is permuted.
Only `r n_perm_over_obs` out of `r length(perm_stats)` permuted mean-log-likelihoods
were larger than the observed mean-log-likelihood
(`r pct_perm_over_obs |> scales::percent(accuracy = 0.1)`),
resulting in a two-sided p-value of $2 \times `r pct_perm_over_obs` = `r 2*pct_perm_over_obs`$.

{{< pagebreak >}}

```{r}
#| fig-cap: |
#|     **Histogram of mean log-likelihoods for data permuted by sex.**
#| label: fig-permute-llik-sex

pval_sex_gp34_v1 |> autoplot() |> print()
```

{{< pagebreak >}}
